---
title: "ORCHIDEE Dedup Framework"
engine: knitr
format:
  html:
    theme: 
      light: flatly
      dark: darkly
    css: assets/orchidee-theme.css
    toc: true
    number-sections: true
    float: true
    code-fold: true
    page-layout: full
    grid:
      sidebar-width: 280px
      body-width: 1450px
      margin-width: 220px
editor: 
  markdown: 
    wrap: 90
editor_options: 
  chunk_output_type: console
---

# Setup

```{r}
#| label: setup_source_bootstrap
#| echo: false
#| warning: false
#| message: false

source(file.path("R", "setup.R"))
```

```{r}
#| label: setup_chunk_options
#| echo: false
#| warning: false
#| message: false

# Chunk options
knitr::opts_chunk$set(
  echo       = TRUE,    # Should blocks with program code be shown in knitted documents?
  eval       = TRUE,    # Should program code be evaluated?
  warning    = FALSE,
  message    = FALSE,
  fig.height = 8,       # Default height for plots.
  fig.width  = 14,       # Default width for plots.
  out.width  = "100%",   # Fill available body width.
  fig.align  = "center" # Default alignment for plots in knitted documents.
)

options(
  scipen = 8,
  digits = getOption("digits"),
  OutDec = ".",
  knitr.kable.NA = ""
 )

```

Ce chunk regroupe les helpers d'export utilisés plus loin dans le notebook. Il prépare le dossier `downloads/`, standardise les noms de fichiers, construit les chemins d'export et fournit les boutons de téléchargement HTML quand le document est rendu.

L'objectif est d'éviter de répéter la même logique d'export dans chaque section de QA ou de résultats.

```{r}
#| label: download_artifact_helpers
download_dir <- orchidee_config$paths$downloads_dir
datatable_digits <- orchidee_config$report$datatable_digits
datatable_filter_default <- orchidee_config$report$datatable_filter_default
datatable_initial_zoom <- orchidee_config$report$datatable_initial_zoom
datatable_zoom_step <- orchidee_config$report$datatable_zoom_step
orchidee_source_required_script("ratb_report_helpers.R", "ratb_report_helpers script")
initialize_ratb_report_helpers(
  download_dir = download_dir,
  datatable_digits = datatable_digits,
  datatable_filter_default = datatable_filter_default,
  datatable_initial_zoom = datatable_initial_zoom,
  datatable_zoom_step = datatable_zoom_step
)
```

## Script sourcing

```{r}
#| label: source_project_scripts
#| echo: false
#| warning: false
#| message: false

orchidee_source_required_script(
  "normalisation_bact.R",
  "normalisation_bact script"
)
orchidee_source_required_script(
  "normalisation_atb.R",
  "normalisation_atb script"
)
orchidee_source_required_script(
  "spares_shared_primitives.R",
  "spares_shared_primitives script"
)
orchidee_source_required_script(
  "phenotype_flag_helpers.R",
  "phenotype_flag_helpers script"
)
orchidee_source_required_script(
  "spares_dedup.R",
  "spares_dedup script"
)
orchidee_source_required_script(
  "completion_helpers.R",
  "completion_helpers script"
)
orchidee_source_required_script(
  "completion_workflow_helpers.R",
  "completion_workflow_helpers script"
)
orchidee_source_required_script(
  "ratb_indicator_helpers.R",
  "ratb_indicator_helpers script"
)
orchidee_source_required_script(
  "ratb_plausibility_qc_helpers.R",
  "ratb_plausibility_qc_helpers script"
)
orchidee_source_required_script(
  "ratb_canonical_runtime_helpers.R",
  "ratb_canonical_runtime_helpers script"
)
orchidee_source_required_script(
  "ratb_hospital_days_helpers.R",
  "ratb_hospital_days_helpers script"
)
orchidee_source_required_script(
  "chu_ratb_scope_adapter.R",
  "CHU native RATB scope adapter"
)
```

## Prebuilt `sir_wide` artifact

This notebook runs in analysis mode and expects a prebuilt `sir_wide` artifact.

Build once (or whenever raw data / mappings / normalization scripts change):

1. `Rscript R/build_sir_wide_artifact.R`
2. `quarto render orchidee_dedup_workflow.qmd`

If artifact validation fails, rerun the builder script above and render again.

```{r}
#| label: load_sir_wide_artifact
#| echo: false

orchidee_source_required_script("helpers.R", "helpers script")

data_dir <- orchidee_config$paths$data_dir
dictionaries_dir <- orchidee_config$paths$dictionaries_dir

sir_wide_artifact_path <- file.path(data_dir, "sir_wide.rds")
sir_wide_meta_path <- file.path(data_dir, "sir_wide_meta.rds")
required_artifacts <- c(sir_wide_artifact_path, sir_wide_meta_path)

if (!all(file.exists(required_artifacts))) {
  missing <- required_artifacts[!file.exists(required_artifacts)]
  stop(
    "Missing sir_wide artifact files: ", paste(missing, collapse = ", "), "\n",
    "Run: Rscript R/build_sir_wide_artifact.R\n",
    "Then rerun: quarto render orchidee_dedup_workflow.qmd",
    call. = FALSE
  )
}

sir_wide_meta <- readRDS(sir_wide_meta_path)
sir_wide <- readRDS(sir_wide_artifact_path)

artifact_validation <- validate_loaded_sir_wide_artifact(
  sir_wide = sir_wide,
  meta = sir_wide_meta
)

sir_wide_artifact_signature <- compute_sir_wide_artifact_signature(
  sir_wide_path = sir_wide_artifact_path,
  sir_wide_meta_path = sir_wide_meta_path,
  meta = sir_wide_meta
)

if (!isTRUE(artifact_validation$ok)) {
  stop(
    "sir_wide artifact is invalid or inconsistent with metadata:\n - ",
    paste(artifact_validation$reasons, collapse = "\n - "),
    "\nRun: Rscript R/build_sir_wide_artifact.R",
    call. = FALSE
  )
}

if (!exists("filtre_atb") && !is.null(sir_wide_meta$filtre_atb)) {
  filtre_atb <- sir_wide_meta$filtre_atb
}
```

### RATB hospitalization scope

The base `sir_wide` artifact remains broad. RATB execution applies the same
CONSORES `CODE_TA` / `CODE_DE` perimeter at two levels: sample `SEJUF` for
microbiology rows, and PMSI episode activity for incidence denominators.

Eligibility contract:

- classify each microbiology and PMSI `SEJUF` with the CONSORES structure reference
- keep a UF when `CODE_TA` is `03` or `20` and `CODE_DE` belongs to a SPARES
  included domain
- keep a microbiology row when its sample `SEJUF` is eligible by this TA/DE rule
- count PMSI episode nights when any PMSI UF is eligible by this TA/DE rule
- keep `PMSISTATUT` as audit context only, not as an inclusion gate
- exclude microbiology rows without an eligible sample `SEJUF` because TA/DE
  eligibility cannot be established

The resulting `sir_wide_ratb_analytic_scope` is used for completion,
deduplication, proportions, and incidence numerators. Eligible hospital nights
are computed separately from PMSI activity.

```{r}
#| label: ratb_scope_cache_control
recompute_ratb_scope <- orchidee_config$cache$recompute_ratb_scope
```

```{r}
#| label: ratb_scope_prepare
#| echo: false

stopifnot(
  exists("build_ratb_scope_tables", mode = "function"),
  exists("build_ratb_downstream_scope_from_canonical_inputs", mode = "function"),
  exists("build_ratb_runtime_payload_from_canonical_inputs", mode = "function"),
  exists("build_ratb_analytic_scope_dataset", mode = "function"),
  exists("build_hospital_days_validation", mode = "function"),
  exists("build_ratb_provisional_perimeter_audit", mode = "function"),
  exists("build_chu_native_ratb_scope_cache_payload", mode = "function")
)

build_ratb_scope_cache_meta <- function(
    sir_wide,
    sir_wide_meta,
    sir_wide_artifact_signature,
    ratb_ref_paths,
    pmsi_input_signature,
    microbiology_scope_policy,
    incidence_denominator_policy
  ) {
  stopifnot(
    is.list(sir_wide_meta),
    is.list(sir_wide_artifact_signature),
    length(ratb_ref_paths) > 0L,
    is.list(pmsi_input_signature)
  )

  scope_script_names <- c(
    "ratb_canonical_runtime_helpers.R",
    "ratb_hospital_days_helpers.R",
    "chu_ratb_scope_adapter.R"
  )
  scope_script_paths <- vapply(
    scope_script_names,
    function(script_name) {
      orchidee_resolve_script_path(
        script_name,
        "script required for RATB scope cache fingerprint"
      )
    },
    character(1)
  )
  scope_script_hashes <- as.list(as.character(unname(tools::md5sum(scope_script_paths))))
  names(scope_script_hashes) <- scope_script_names
  reference_hashes <- as.list(as.character(unname(tools::md5sum(ratb_ref_paths))))
  names(reference_hashes) <- names(ratb_ref_paths)

  payload <- list(
    sir_wide_n_rows = nrow(sir_wide),
    sir_wide_n_eltid = dplyr::n_distinct(sir_wide$ELTID),
    artifact_created_at = sir_wide_meta$created_at,
    sir_wide_artifact_signature = sir_wide_artifact_signature,
    microbiology_scope_policy = microbiology_scope_policy,
    incidence_denominator_policy = incidence_denominator_policy,
    scope_script_hashes = scope_script_hashes,
    pmsi_input_signature = pmsi_input_signature,
    reference_hashes = reference_hashes
  )

  tmp_file <- tempfile(pattern = "ratb_scope_fingerprint_", fileext = ".rds")
  on.exit(unlink(tmp_file), add = TRUE)
  saveRDS(payload, tmp_file, version = 2)
  fingerprint <- as.character(unname(tools::md5sum(tmp_file)))

  list(
    fingerprint = fingerprint,
    sir_wide_n_rows = nrow(sir_wide),
    sir_wide_n_eltid = dplyr::n_distinct(sir_wide$ELTID),
    artifact_created_at = sir_wide_meta$created_at,
    sir_wide_artifact_signature = sir_wide_artifact_signature,
    microbiology_scope_policy = microbiology_scope_policy,
    incidence_denominator_policy = incidence_denominator_policy,
    scope_script_hashes = scope_script_hashes,
    pmsi_input_signature = pmsi_input_signature,
    reference_hashes = reference_hashes,
    created_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
  )
}

ratb_file_signature <- function(path) {
  info <- file.info(path)
  list(
    size_bytes = unname(as.numeric(info$size)),
    md5 = as.character(unname(tools::md5sum(path)))
  )
}

ratb_ref_paths <- vapply(
  c(
    ref_uf = file.path(orchidee_config$paths$ref_dir, "ref_uf.txt"),
    ref_um = file.path(orchidee_config$paths$ref_dir, "ref_um.txt"),
    ref_uf2um = file.path(orchidee_config$paths$ref_dir, "ref_uf2um.txt"),
    consores_structure = orchidee_config$paths$consores_structure_path,
    consores_codes_ta = orchidee_config$paths$consores_codes_ta_path,
    consores_codes_de = orchidee_config$paths$consores_codes_de_path
  ),
  function(x) resolve_existing_path(c(x), what = paste0("reference ", x)),
  character(1)
)
ratb_pmsi_path_candidates <- c("pmsi", file.path(data_dir, "pmsi"))
ratb_pmsi_input_path <- resolve_existing_path(
  ratb_pmsi_path_candidates,
  what = "pmsi raw input"
)

ratb_scope_cache_candidates <- list(
  list(
    payload = file.path("data", "ratb_scope_cache"),
    meta = file.path("data", "ratb_scope_cache_meta")
  )
)
expected_ratb_scope_payload_names <- c(
  "sample_scope_reference",
  "denominator_bundle",
  "ratb_scope_join_audit",
  "ratb_scope_exclusion_summary",
  "hospital_stays_raw",
  "hospital_stays_validated",
  "hospital_stay_validation_summary",
  "hospital_days_year_split",
  "hospital_days_year_summary",
  "ratb_perimeter_rules",
  "ratb_uf_ta_de_reference",
  "ratb_episode_scope_audit",
  "ratb_episode_exclusion_summary",
  "hospital_days_year_split_provisional",
  "hospital_days_year_summary_provisional",
  "ratb_numerator_scope_impact_audit"
)

ratb_scope_cache_script_hashes_match <- function(loaded_meta, current_meta, script_names) {
  is.list(loaded_meta$scope_script_hashes) &&
    is.list(current_meta$scope_script_hashes) &&
    all(vapply(
      script_names,
      function(script_name) {
        identical(
          loaded_meta$scope_script_hashes[[script_name]],
          current_meta$scope_script_hashes[[script_name]]
        )
      },
      logical(1)
    ))
}

ratb_scope_cache_canonical_inputs_match <- function(loaded_meta, current_meta) {
  fields <- c(
    "sir_wide_n_rows",
    "sir_wide_n_eltid",
    "artifact_created_at",
    "sir_wide_artifact_signature",
    "microbiology_scope_policy",
    "incidence_denominator_policy",
    "pmsi_input_signature",
    "reference_hashes"
  )
  producer_script_names <- c(
    "ratb_hospital_days_helpers.R",
    "chu_ratb_scope_adapter.R"
  )

  is.list(loaded_meta) &&
    all(vapply(
      fields,
      function(field) identical(loaded_meta[[field]], current_meta[[field]]),
      logical(1)
    )) &&
    ratb_scope_cache_script_hashes_match(
      loaded_meta = loaded_meta,
      current_meta = current_meta,
      script_names = producer_script_names
    )
}

ratb_scope_cache_source <- NA_character_
ratb_scope_cache_payload <- NULL
ratb_scope_cache_meta <- NULL
ratb_scope_cache_decision <- NA_character_
ratb_scope_cache_refresh_needed <- FALSE

ratb_scope_cache_meta_current <- build_ratb_scope_cache_meta(
  sir_wide = sir_wide,
  sir_wide_meta = sir_wide_meta,
  sir_wide_artifact_signature = sir_wide_artifact_signature,
  ratb_ref_paths = ratb_ref_paths,
  pmsi_input_signature = ratb_file_signature(ratb_pmsi_input_path),
  microbiology_scope_policy = orchidee_config$ratb$microbiology_scope_policy,
  incidence_denominator_policy = orchidee_config$ratb$incidence_denominator_policy
)

if (!isTRUE(recompute_ratb_scope)) {
  for (candidate in ratb_scope_cache_candidates) {
    cache_paths <- unlist(candidate, use.names = FALSE)
    if (!all(file.exists(cache_paths))) {
      next
    }

    loaded_meta <- tryCatch(readRDS(candidate$meta), error = function(e) NULL)
    if (!is.list(loaded_meta) || is.null(loaded_meta$fingerprint)) {
      next
    }
    cache_fingerprint_matches <- identical(
      as.character(loaded_meta$fingerprint),
      ratb_scope_cache_meta_current$fingerprint
    )
    cache_canonical_inputs_match <- cache_fingerprint_matches ||
      ratb_scope_cache_canonical_inputs_match(
        loaded_meta = loaded_meta,
        current_meta = ratb_scope_cache_meta_current
      )
    if (!cache_canonical_inputs_match) {
      next
    }

    loaded_payload <- tryCatch(readRDS(candidate$payload), error = function(e) NULL)
    if (!is.list(loaded_payload) || !all(expected_ratb_scope_payload_names %in% names(loaded_payload))) {
      next
    }
    if (!is.data.frame(loaded_payload$sample_scope_reference)) {
      next
    }
    if (!is.list(loaded_payload$denominator_bundle) || !all(c(
      "hospital_days_year_summary",
      "hospital_days_year_summary_provisional"
    ) %in% names(loaded_payload$denominator_bundle))) {
      next
    }
    loaded_denominator_provisional <-
      loaded_payload$denominator_bundle$hospital_days_year_summary_provisional
    if (!is.data.frame(loaded_denominator_provisional) ||
        !"hospital_nights_provisional" %in% names(loaded_denominator_provisional)) {
      next
    }
    loaded_scope_base <- loaded_payload$sir_wide_ratb_scope_base
    if (!is.data.frame(loaded_scope_base) && is.data.frame(loaded_payload$sir_wide_ratb_scope)) {
      loaded_scope_base <- loaded_payload$sir_wide_ratb_scope
    }
    if (!is.data.frame(loaded_scope_base)) {
      next
    }
    if (nrow(loaded_scope_base) > nrow(sir_wide)) {
      next
    }
    if (!"ratb_scope_status" %in% names(loaded_scope_base)) {
      next
    }
    if (!all(loaded_scope_base$ratb_scope_status %in% c(
      "eligible_hospitalization",
      "excluded_external",
      "mixed_status",
      "no_usable_status",
      "no_pmsi_match"
    ))) {
      next
    }
    if (!all(loaded_payload$hospital_stays_validated$validation_status == "validated")) {
      next
    }
    if (!all(!loaded_payload$hospital_stays_validated$missing_bounds)) {
      next
    }
    if (!all(!loaded_payload$hospital_stays_validated$negative_elapsed)) {
      next
    }
    if (!all(loaded_denominator_provisional$hospital_nights_provisional >= 0)) {
      next
    }

    ratb_scope_cache_payload <- loaded_payload
    ratb_scope_cache_meta <- loaded_meta
    ratb_scope_cache_source <- candidate$payload
    ratb_scope_cache_refresh_needed <- !cache_fingerprint_matches
    break
  }
} else {
  ratb_scope_cache_source <- "forced_recompute"
}

if (!isTRUE(recompute_ratb_scope) && !is.null(ratb_scope_cache_payload)) {
  ratb_scope_cache_payload <- build_ratb_runtime_payload_from_canonical_inputs(
    payload = ratb_scope_cache_payload,
    sir_wide = sir_wide
  )
  if (isTRUE(ratb_scope_cache_refresh_needed)) {
    ratb_scope_cache_meta <- ratb_scope_cache_meta_current
    saveRDS(ratb_scope_cache_payload, file.path("data", "ratb_scope_cache"))
    saveRDS(ratb_scope_cache_meta, file.path("data", "ratb_scope_cache_meta"))
  }
  ratb_scope_cache_decision <- if (isTRUE(ratb_scope_cache_refresh_needed)) {
    "loaded_refreshed"
  } else {
    "loaded"
  }
  message("Loaded RATB scope artifacts from ", ratb_scope_cache_source)
} else {
  if (isTRUE(recompute_ratb_scope)) {
    message("Force recompute requested; building RATB scope artifacts.")
  } else {
    message("No valid RATB scope cache found; building RATB scope artifacts.")
  }

  native_scope_cache <- build_chu_native_ratb_scope_cache_payload(
    sir_wide = sir_wide,
    pmsi_path_candidates = ratb_pmsi_path_candidates,
    structure_path = orchidee_config$paths$consores_structure_path,
    codes_ta_path = orchidee_config$paths$consores_codes_ta_path,
    codes_de_path = orchidee_config$paths$consores_codes_de_path,
    ref_dir = orchidee_config$paths$ref_dir
  )
  ratb_scope_cache_payload <- native_scope_cache$payload
  ratb_scope_cache_payload <- build_ratb_runtime_payload_from_canonical_inputs(
    payload = ratb_scope_cache_payload,
    sir_wide = sir_wide
  )
  ratb_scope_cache_meta <- ratb_scope_cache_meta_current

  saveRDS(ratb_scope_cache_payload, file.path("data", "ratb_scope_cache"))
  saveRDS(ratb_scope_cache_meta, file.path("data", "ratb_scope_cache_meta"))
  ratb_scope_cache_source <- "recomputed"
  ratb_scope_cache_decision <- "recomputed"
  message("Saved RATB scope artifacts to: data/ratb_scope_cache, data/ratb_scope_cache_meta")

  rm(native_scope_cache)
}

list2env(ratb_scope_cache_payload, envir = environment())

ratb_runtime_validation <- validate_ratb_canonical_runtime_inputs(
  runtime_inputs = ratb_scope_cache_payload,
  sir_wide = sir_wide
)
if (!isTRUE(ratb_runtime_validation$ok)) {
  stop(
    "Canonical RATB runtime inputs are invalid:\n - ",
    paste(ratb_runtime_validation$errors, collapse = "\n - "),
    call. = FALSE
  )
}

stopifnot(
  all(
    !ratb_episode_scope_audit$included_in_provisional_perimeter |
      ratb_episode_scope_audit$n_ta_de_eligible_uf > 0L
  ),
  all(hospital_stays_validated$validation_status == "validated"),
  all(!hospital_stays_validated$missing_bounds),
  all(!hospital_stays_validated$negative_elapsed),
  all(hospital_days_year_summary_provisional$hospital_nights_provisional >= 0)
)

cat("recompute_ratb_scope: ", isTRUE(recompute_ratb_scope), "\n", sep = "")
cat("ratb_scope_cache_source: ", ratb_scope_cache_source, "\n", sep = "")
cat("ratb_scope_cache_decision: ", ratb_scope_cache_decision, "\n", sep = "")
```

### RATB biological plausibility QC

The RATB analytical scope is filtered before completion and deduplication so
biologically implausible susceptibility patterns cannot be propagated by
completion or retained as dedup representatives.

```{r}
#| label: ratb_plausibility_qc
ratb_plausibility_qc <- build_ratb_plausibility_qc(
  sir_df = sir_wide_ratb_analytic_scope,
  species_regex_map_path = file.path(dictionaries_dir, "species_regex_map.csv")
)

sir_wide_ratb_analytic_scope_qc <- ratb_plausibility_qc$data
ratb_plausibility_qc_excluded <- ratb_plausibility_qc$excluded
ratb_plausibility_qc_summary <- ratb_plausibility_qc$summary
ratb_plausibility_qc_unavailable_rules <- ratb_plausibility_qc$unavailable_rules

stopifnot(
  nrow(sir_wide_ratb_analytic_scope_qc) + nrow(ratb_plausibility_qc_excluded) == nrow(sir_wide_ratb_analytic_scope),
  all(ratb_plausibility_qc_excluded$qc_any_plausibility_exclusion)
)

ratb_plausibility_qc_summary
```

```{r}
#| label: ratb_scope_exclusion_summary_table
show_audit_table(
  data = ratb_scope_exclusion_summary,
  file_stem = "ratb_scope_exclusion_summary",
  caption = "RATB hospitalization scope summary",
  page_length = 10,
  scroll_y = "260px",
  digits = datatable_digits
)
```

```{r}
#| label: ratb_scope_join_audit_table
show_audit_table(
  data = ratb_scope_join_audit %>%
    select(
      PATID, EVTID, n_micro_rows, n_eltid, ratb_scope_status,
      included_in_ratb_scope, pmsi_status_values, n_status_non_missing,
      n_distinct_status, evtid_multi_pat
    ),
  file_stem = "ratb_scope_join_audit",
  caption = "RATB scope join audit (PATID, EVTID)",
  page_length = 20,
  scroll_y = "360px"
)
```

## Quickstart for colleagues

This notebook can be used in two ways:

- **Analyst mode**: render and inspect outputs, with cache-first behavior.
- **Maintainer mode**: update inputs/rules/scripts, then rebuild artifacts and rerender.

### Step-1 setup (cold start)

1. Build base artifact: `Rscript R/build_sir_wide_artifact.R`
2. Render notebook: `quarto render orchidee_dedup_workflow.qmd`

### Step-2 routine (warm run)

Use plain render first:

- `quarto render orchidee_dedup_workflow.qmd`

Then check cache diagnostics in output:

- `ratb_scope_cache_decision` (`loaded`, `loaded_refreshed`, or `recomputed`)
- `completion_cache_decision` (`loaded` or `recomputed`)
- `dedup_cache_decision` (`loaded` or `recomputed`)

### Run modes (what to recompute)

- **Presentation-only changes** (text, themes, plots): just render.
- **RATB scope / denominator input or producer-logic changes** (`data/pmsi`, `ratb_hospital_days_helpers.R`, `chu_ratb_scope_adapter.R`, `ref/ref_uf*.txt`, `ref/ref_um.txt`, CONSORES TA/DE reference files): render the notebook. The scope cache fingerprint includes PMSI content, producer helper hashes and reference hashes; set `orchidee_config$cache$recompute_ratb_scope <- TRUE` only when you need to force a rebuild manually.
- **Completion logic changes** (`completion_*` helpers/rules): set `orchidee_config$cache$recompute_completion <- TRUE` in `config/pipeline.R`, render once, then revert to `FALSE`.
- **Dedup logic changes** (`spares_dedup.R` / scope defs): set `orchidee_config$cache$recompute_dedup <- TRUE` in `config/pipeline.R`, render once, then revert to `FALSE`.
- **Microbiology artifact changes**: rerun builder script, then render (downstream caches use the canonical `sir_wide` artifact signature and should auto-invalidate where relevant).

- **Helper script changes**: rerender the notebook. If the cached canonical
  inputs still match the current `sir_wide`, policies and references, the
  notebook can refresh runtime scope objects from those inputs. Otherwise a
  native RATB scope recompute is expected.
## Control panel (edit points)

This notebook is intentionally parameterized through `config/pipeline.R` plus a
few local derived objects.

| What you want to tune | Variable(s) | Where |
|---|---|---|
| Datatable decimals / filters / zoom UI | `orchidee_config$report` | `config/pipeline.R` |
| RATB scope / denominator recompute policy | `orchidee_config$cache$recompute_ratb_scope` | `config/pipeline.R` |
| Completion recompute policy | `orchidee_config$cache$recompute_completion` | `config/pipeline.R` |
| Dedup recompute policy | `orchidee_config$cache$recompute_dedup` | `config/pipeline.R` |
| RATB indicator spec and display thresholds | `orchidee_config$paths$ratb_indicator_spec_path`, `orchidee_config$ratb` | `config/pipeline.R` |

## Cache and artifact contract

| Artifact | Produced by | Consumed by | Refresh trigger |
|---|---|---|---|
| `data/sir_wide.rds`, `data/sir_wide_meta.rds` | `R/build_sir_wide_artifact.R` | `load_sir_wide_artifact` | Raw inputs, dictionaries, normalization script changes |
| `data/ratb_scope_cache`, `data/ratb_scope_cache_meta` | RATB scope recompute path | hospitalization scope + hospital-days sections | RATB scope fingerprint mismatch or `orchidee_config$cache$recompute_ratb_scope <- TRUE` |
| `data/completion_datasets`, `data/completion_logs`, `data/raw_row_log`, `data/completion_cache_meta` | completion recompute path | completion + dedup sections | completion fingerprint mismatch or `orchidee_config$cache$recompute_completion <- TRUE` |
| `data/dedup_results`, `data/dedup_cache_meta` | dedup recompute path | dedup + resistance sections | dedup/completion fingerprint mismatch or `orchidee_config$cache$recompute_dedup <- TRUE` |
| `downloads/*.pdf`, `downloads/*.xlsx`, `downloads/*.png` | helper exporters | reader download buttons | regenerated at render |

Notes:

- Cache system is custom and **independent from knitr cache**.
- Cache read order is root first, then `data/` fallback where configured.
- Cache fingerprints include canonical artifact signatures and selected
  helper-script/reference hashes. Runtime-helper-only changes can reuse and
  refresh matching canonical inputs; PMSI, reference or producer-helper
  changes force a native RATB scope recompute.

## Troubleshooting quick fixes

- **Error: missing `sir_wide` artifact files**
  - Run: `Rscript R/build_sir_wide_artifact.R`
  - Rerender notebook.

- **Error: invalid or inconsistent `sir_wide` artifact**
  - Rebuild `sir_wide` artifact and rerender.
  - Confirm dictionary/script paths are the expected ones.

- **Render succeeds but outputs look unchanged**
  - Check `completion_cache_decision` / `dedup_cache_decision` in report.
  - If needed, force recompute (`recompute_completion` and/or `recompute_dedup`) for one render cycle.

- **Heatmaps too dense after scope expansion**
  - Increase baseline heatmap sizing parameters in `resistance_heatmap_controls_pairwise`.
  - Raise `res_heatmap_min_n` or temporarily set `res_heatmap_pairwise_show_labels <- FALSE`.

- **`quarto` not found in shell**
  - Use terminal where Quarto is installed/in PATH (or full executable path).

# Methods
Base artifact: `sir_wide` (broad microbiology build, 1 row = 1 strain-level antibiogram record).

Operational RATB inputs:
- `sir_wide_ratb_scope`, the broad microbiology table annotated with
  sample-level TA/DE scope and CHU join audit context;
- `sir_wide_ratb_analytic_scope`, the sample-level TA/DE microbiology perimeter used for completion, deduplication, proportions, and incidence numerators;
- `hospital_days_year_summary_provisional`, the PMSI TA/DE denominator built independently from microbiology rows.

Reader roadmap:
- derive and audit the sample-level microbiology TA/DE scope from the broad artifact;
- derive and audit the PMSI TA/DE denominator independently from microbiology rows;
- build and validate completion datasets under multiple grouping strategies;
- run SPARES-aligned dedup on all datasets on the microbiology analytical base;
- audit structural drift, class partition drift, representative identity drift, and C3G indicator drift;
- validate and document the PMSI-based `hospital_days` denominator layer on the same analytical perimeter;
- synthesize a final go/no-go decision for the default operational pipeline.

## Decision objective

Beside "simply" reproducing SPARES, there's an important methodological question: whether completion changes downstream dedup outcomes enough to justify added complexity across heterogeneous hospital reporting practices. We therefore run multiple completion strategies, keep dedup logic constant, and compare safety + impact before deciding what is worth operationalizing.

## Completion strategy matrix

| dataset | group_keys | max_days | what changes | what stays invariant |
|---|---|---:|---|---|
| `sir_wide_raw` | none | NA | no completion transfer | row count, `ELTID` count, original ATB values |
| `completion_date` | `PATID, EVTID, DATEPRELEV, naturepvt_norm, bact_norm` | `Inf` | within-group class-aware S/R fill | non-collapsing rows, no non-NA overwrite, no ZIT fill |
| `completion_evtid` | `PATID, EVTID, naturepvt_norm, bact_norm` | `Inf` | same rules on patient-event scope | same invariants as above |
| `completion_patid_30d` | `PATID, naturepvt_norm, bact_norm` | `30` | same rules with patient-level time guard | same invariants + 30-day constraint |
| `completion_patid_year` | `PATID, completion_evt_start_year, naturepvt_norm, bact_norm` | `Inf` | same rules within patient-level stay-start year groups | same invariants as above |

`ATB` cross-resistance inference / microbiologist-driven completion rules are intentionally
out of scope for this notebook version.

## Phenotype integration

BLSE / carbapenemase parsing is performed upstream in `build_sir_wide_artifact.R`,
before the `LBLRES == "SIR"` reduction. This is intentional: phenotype-bearing
rows exist only in the broader `bact22_24` source and would otherwise be
dropped before completion / dedup ever sees them.

The artifact-level collapse key is aligned with the future `sir_wide` row grain:
`(PATID, EVTID, ELTID, DATEPRELEV, souche_id, bact_norm, naturepvt_norm)`.

Before phenotype parsing, the artifact builder excludes whole `ELTID`s carrying
one of the three locked screening `TYPEANA` codes (`BGBLSE_R.BGBLSE_R2`,
`BGCARBA_R.BGCARBA_R2`, `BGABMR_R.BGABMR_R2`). This makes the screening exclusion
a shared microbiology perimeter rule for all indicators, not a phenotype-only
exception.

`phenotype_flag_helpers.R` then classifies the remaining phenotype-bearing rows
from `(LBLANA, LBLRES, STRRES)` using exact-match rules first and targeted text
fallbacks second. Ambiguous free-text formulations such as `cote blse` /
`cote carba` are intentionally retained as `unknown`.

The artifact exposes:

- `blse_status_row`, `carbapenemase_status_row`
- `blse_flag`, `carbapenemase_flag`

Completion and dedup do not operate directly on the public flags. They derive
temporary S/R proxy columns:

- `.pheno_blse_sr`
- `.pheno_carba_sr`

Encoding:

- `positive -> R`
- `negative` / `unknown` / `no_signal -> S`

Implications:

- completion may propagate phenotype only within a compatible class;
- a positive phenotype disagreement with any non-positive phenotype status creates a hard class split;
- for `Staphylococcus aureus`, the methicillin indicator is resolved downstream by ordered surrogate logic (`cefoxitine`, then `oxacilline`); current raw data does not appear to contain `cefoxitine` for `S. aureus`, so this is currently numerically equivalent to `oxacilline` alone.
- for SPARES deduplication, absence of a BLSE / carbapenemase result is interpreted as absence of that phenotype.

After completion and representative selection, phenotype is summarised again at
the retained-class level. Final outputs expose:

- `blse_status_final`, `carbapenemase_status_final`
- `blse_flag`, `carbapenemase_flag`

Validation targets for this integration:

- unresolved `(LBLANA, LBLRES, STRRES)` tails remain auditable;
- no retained class may contain both positive and non-positive status for the same
  phenotype;
- phenotype indicators should no longer be gated as
  `pending_phenotype_flag` once the final dedup outputs carry the public flags.

## Dedup rulebook

### Class compatibility rule 

Within each dedup episode, records are partitioned into phenotype-aware compatibility classes with strict
pairwise compatibility:

- A pair is incompatible if at least one major discordance `S<->R` exists on informative overlap.
- `NA` and `ZIT` are treated as non-informative for discordance checks.
- Class assignment uses deterministic greedy first-fit and must preserve all-pairs non-discordance.

### Why compatibility is non-transitive

Compatibility is symmetric but can be non-transitive when `NA`/`ZIT` bridge otherwise
discordant profiles.

Example (single ATB):

| Document | ceftriaxone |
|----------|-------------|
| A        | R           |
| B        | NA          |
| C        | S           |

Then:

- `A ~ B` (no informative conflicting overlap)
- `B ~ C` (no informative conflicting overlap)
- `A !~ C` (discordance `R<->S`)

So `A ~ B` and `B ~ C` do not imply `A ~ C`. `B` acts as a bridge compatible with both
sides, while the sides remain mutually incompatible.

Should `B` belong to class `A` or class `C`?

Order-sensitivity example:

- If processing order is `A`, `C`, then `B`:
  - `A` opens class 1.
  - `C` opens class 2 (incompatible with `A`).
  - `B` is compatible with both classes.
- With first-fit, `B` joins the first compatible class encountered.
- If class order is reversed, `B` can be assigned to the other class.

Both outcomes remain valid (no within-class conflict), but the partition can differ.
This sensitivity can be measured explicitly (e.g., normal vs reversed order).

When transitivity tends to hold:

- panels are fully observed (no `NA`/`ZIT`)
- compared records share the same tested-ATB mask

### Representative selection rule

Exactly one representative record is retained per compatibility class. Deterministic sort
priority:

1. dataset-specific completeness:
   - `sir_wide_raw`: pre-completion completeness (`nb_resultats`)
   - completion datasets: post-completion completeness (`nb_resultats_post`)
2. `evt_order`
3. `elt_order`
4. tie-break on `ELTID`, then `souche_id`

## Contracts and invariants

Each invariant is defined once here and checked later in executable diagnostics.

- Completion is non-collapsing: one output row per input row.
- Structural invariants: `nrow(dataset) == nrow(sir_wide_raw)` and
  `n_distinct(ELTID)` unchanged.
- Value safety: no mutation of originally non-`NA` ATB values (including non-`NA -> NA`).
- Value domain safety: no completion-generated `ZIT` from missing values.
- Monotonic completeness: `nb_resultats_post >= nb_resultats` for every row.
- Conflict safety: no transfer across major `S<->R` discordance.
- Time safety: `completion_patid_30d` never uses an effective fill source beyond 30 days.
- Schema safety: completion datasets/logs expose the expected column contracts.
- Dedup audit invariant: `n_within_class_discord_pairs == 0` across scopes.

## Canonical dictionary (objects + metrics + invariants)

### Objects

- `completion_datasets`: named list with `sir_wide_raw`, `completion_date`,
  `completion_evtid`, `completion_patid_30d`, `completion_patid_year`.
- `completion_logs`: named list with per-strategy `group_log` and `row_log`.
- `completion_cache_meta`: cache fingerprint metadata used to validate completion cache freshness.
- `dedup_results` (cache artifact): nested list of dedup outputs by `(dataset, scope)`.
- `dedup_cache_meta`: cache fingerprint metadata used to validate dedup cache freshness.
- `completion_checks`: cross-dataset safety gate table.
- `strategy_impact_summary`: per-strategy impact synthesis after completion.
- `strategy_drift_summary`: pairwise row-equality drift across strategies.
- `conflict_consistency`: helper-vs-matrix discordance consistency check (computed once and reused by diagnostics).
- `dedup_summary`: dedup outcomes per `(dataset, scope)` on SPARES-aligned PATID-year scopes.
- `dedup_impact_summary`: deltas vs raw dedup baseline, including representative identity drift.
- `partition_drift_summary`: episode-level class partition drift counts vs raw by scope.
- `partition_drift_episodes`: changed episodes (partition and/or class-count drift) for review.
- `swap_decomposition_summary`: representative swaps split by partition-changed vs same-partition episodes.
- `swap_decomposition_episodes`: episode-level swap drill-down with added/removed counts and partition flags.
- `c3g_indicator_panel` (panel output): per `(dataset, scope, dedup_year, naturepvt_norm, bact_norm)` C3G counts/rates from dedup outputs.
- `c3g_indicator_impact` and `c3g_indicator_impact_summary` (drift outputs): C3G panel deltas and compact synthesis vs raw baseline, including denominator comparability counts for percent metrics.
- `c3g_rate_global_dedup` and `c3g_rate_by_type_dedup` (panel outputs): operational C3G rates (`n_resistant / n_tested`) across all dedup datasets.
- `c3g_rate_global_drift_vs_raw` and `c3g_rate_by_type_drift_vs_raw` (drift outputs): C3G rate deltas vs raw baseline.
- `c3g_rate_global_raw_dedup`, `c3g_rate_by_type_raw_dedup` (compatibility aliases): raw-only extracts retained for older downstream snippets.

### Completion log metrics

- `n_pairs_unique`: `n_rows * (n_rows - 1) / 2` (unordered pairs in class/group).
- `n_pairs_checked`: directional pair checks over forward/backward passes.
- `n_pairs_conflict`: directional checks blocked by `S<->R` discordance.
- `n_cells_filled_total`: total transferred cells in class/group.
- `n_rows_filled`: rows with `n_cells_filled > 0`.
- `n_iterations`: pass cycles before convergence.
- `max_gap_used_days`: max effective source gap per row.

### Completion safety metrics (`completion_checks`)

- `same_n_rows_as_raw`
- `same_n_eltid_as_raw`
- `non_na_overwrite_count` (must be `0`)
- `non_na_to_na_loss_count` (must be `0`)
- `filled_with_zit_count` (must be `0`)
- `monotonic_nb_resultats` (must be `TRUE`)
- `pct_rows_improved` (`100 * mean(nb_resultats_post > nb_resultats)`)

### Diagnostic summary outputs

- `completion_gate`: compact projection of `completion_checks`.
- `group_kpis`: aggregate group-log activity by strategy.
- `top_fills`: highest-impact completed rows in trusted sample types.
- `completion_diagnostics$conflict_consistency`: compatibility alias/view of canonical `conflict_consistency`.
- `eltid_structural_distribution`, `eltid_structural_outliers`: document-level structure baseline.
- `evtid_structural_distribution`, `patid_structural_distribution`: event-level and patient-level structure baselines.
- Canonical dedup/C3G diagnostics are listed once in `### Objects` above (panel outputs, drift outputs, and compatibility aliases).

### Dedup summary metrics

Core dedup outputs are reported per `(dataset, scope)`, where `scope` is `by_type` or `global`.

- `n_input_rows`: number of rows entering dedup for the given dataset/scope (before representative selection).
- `n_retained_rows`: number of rows kept after dedup (one representative per compatibility class).
- `n_episodes`: number of dedup groups/episodes evaluated in that scope.
- `pct_episodes_multi_class`: `100 * mean(has_multiple_classes)` across episodes.  
  Interpretation: share of episodes split into more than one compatibility class after applying the antibiogram and phenotype rules.
- `order_sensitive_rate_pct`: `100 * mean(order_sensitive)` across episodes with audit info.  
  Interpretation: share of episodes where first-fit class assignment depends on row order (non-transitivity risk).
- `n_within_class_discord_pairs`: sum of within-class major discordant pairs (`S<->R`) across audited episodes.  
  Safety invariant: must stay `0`.

### Dedup impact metrics (`dedup_impact_summary`, vs `sir_wide_raw`)

- `delta_retained_rows`: `n_retained_rows - raw_n_retained_rows` (same scope baseline).
- `delta_pct_episodes_multi_class`: drift in multi-class episode rate (percentage points).
- `delta_order_sensitive_rate_pct`: drift in order-sensitivity rate (percentage points).
- `n_within_class_discord_pairs`: carried as a safety control column (must stay `0`).
- `n_rep_removed_vs_raw`: representatives present in raw dedup output but absent in current dataset output.
- `n_rep_added_vs_raw`: representatives absent in raw dedup output but present in current dataset output.
- `n_rep_swaps_vs_raw`: total representative identity drift, defined as
  `n_rep_removed_vs_raw + n_rep_added_vs_raw`.
- `pct_rep_swaps_vs_raw`: `100 * n_rep_swaps_vs_raw / n_retained_rows`.

Quick reading guide:

- `delta_retained_rows ~= 0` with non-zero `n_rep_swaps_vs_raw` means stable totals but changed representative identity.
- Large absolute `delta_order_sensitive_rate_pct` means order sensitivity changed materially vs raw baseline.

## Minimal implementation map

1. Shared primitives/helpers:
- `is_sr`
- `.spares_sr_conflict_pair(...)` from `spares_shared_primitives.R`
- `pair_gap_days_ok`
- `transfer_sr`
- `complete_group_bidirectional`

2. Generic completion wrapper:
- `run_completion_strategy(df, group_keys, atb_cols, strategy_name, max_days, zit_values)`

3. Strategy instantiation:
- `completion_date`: `group_by(PATID, EVTID, DATEPRELEV, naturepvt_norm, bact_norm)`, `max_days = Inf`
- `completion_evtid`: `group_by(PATID, EVTID, naturepvt_norm, bact_norm)`, `max_days = Inf`
- `completion_patid_30d`: `group_by(PATID, naturepvt_norm, bact_norm)`, `max_days = 30`
- `completion_patid_year`: `group_by(PATID, completion_evt_start_year, naturepvt_norm, bact_norm)`, `max_days = Inf`

4. Impact and dedup diagnostics:
- assemble completion impact from `completion_checks` + `completion_logs`
- compare dedup deltas against `sir_wide_raw` by reporting scope

## Validation targets

- row count unchanged vs raw
- `ELTID` count unchanged vs raw
- no overwrite of originally non-`NA` ATB values
- no completion-generated `ZIT` from missing cells
- `nb_resultats_post >= nb_resultats` for all rows
- no transfer across `S<->R` conflicts
- 30-day guard enforced for `completion_patid_30d`
- completion schema contract aligned across all datasets/logs
- dedup audit invariant `n_within_class_discord_pairs == 0`

# Exploratory diagnostics

## Tile map distribution nb_resultats per group

```{r}
#| label: plot_nb_resultats_tile_prep

df_plot <- sir_wide %>%
  filter(naturepvt_norm %in% c("hemoculture", "urines")) %>%
  mutate(DATEPRELEV = as.Date(DATEPRELEV), year = year(DATEPRELEV)) %>%
  filter(!is.na(year), !is.na(nb_resultats), !is.na(bact_norm))

df_tile <- df_plot %>%
  count(naturepvt_norm, bact_norm, nb_resultats, name = "n", .drop = FALSE) %>%
  group_by(naturepvt_norm, bact_norm) %>%
  mutate(pct = n / sum(n, na.rm = TRUE)) %>%
  ungroup()

```

### Tile map

```{r}
#| label: plot_nb_resultats_tile
p_nb_resultats_tile <- ggplot(df_tile, aes(x = nb_resultats, y = bact_norm, fill = pct)) +
  geom_tile() +
  geom_text(aes(label = ifelse(n == 0, "", n), color = pct > 0.25), size = 2.8) +
  scale_color_manual(values = c(`TRUE` = "white", `FALSE` = "black"), guide = "none") +
  facet_wrap(~ naturepvt_norm, scales = "fixed", drop = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_fill_viridis_c(
    option = "viridis",
    trans = "sqrt",
    direction = -1,
    breaks = c(0, 0.02, 0.05, 0.10, 0.25, 0.50, 0.75, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(x = "nb_resultats", y = "bact_norm", fill = "% (trans = 'sqrt')") +
  theme_minimal()

print(p_nb_resultats_tile)
tile_pdf_path <- save_plot_pdf(p_nb_resultats_tile, "plot_nb_resultats_tile")
show_download_button(
  tile_pdf_path,
  paste0("Download plot as PDF (", basename(tile_pdf_path), ")")
)

```

## Dot map distribution nb_resultats per group

```{r}
#| label: plot_nb_resultats_dot_prep
df_nonzero <- df_tile %>% filter(n > 0)

```

### Dot map

```{r}
#| label: plot_nb_resultats_dot
p_nb_resultats_dot <- ggplot(df_nonzero, aes(x = nb_resultats, y = bact_norm)) +
  geom_point(aes(size = pct, color = pct), alpha = 0.95) +
  geom_text(
    data = df_nonzero %>% filter(pct > 0.25),
    aes(label = n),
    color = "white",
    size = 2.6
  ) +
  geom_text(
    data = df_nonzero %>% filter(pct <= 0.25),
    aes(label = n),
    color = "black",
    size = 2.6
  ) +
  facet_wrap(~ naturepvt_norm, scales = "fixed", drop = FALSE) +
  scale_x_continuous(breaks = scales::pretty_breaks()) +
  scale_colour_viridis_c(
    option = "viridis",
    direction = -1,
    trans = "sqrt",
    breaks = c(0, 0.02, 0.05, 0.10, 0.25, 0.50, 0.75, 1),
    labels = scales::percent_format(accuracy = 1)
  ) +
  scale_size_continuous(range = c(2.5, 8)) +
  theme_minimal() +
  labs(x = "nb_resultats", y = "bact_norm", color = "%", size = "%")

print(p_nb_resultats_dot)
dot_pdf_path <- save_plot_pdf(p_nb_resultats_dot, "plot_nb_resultats_dot")
show_download_button(
  dot_pdf_path,
  paste0("Download plot as PDF (", basename(dot_pdf_path), ")")
)
```

# Completion

## Execution

### Preparation

#### Build raw completion baseline

```{r}
#| label: completion_prepare_raw_baseline
stopifnot(exists("sir_wide_ratb_analytic_scope_qc"), is.data.frame(sir_wide_ratb_analytic_scope_qc))

sir_wide_raw <- sir_wide_ratb_analytic_scope_qc %>%
  mutate(.date_prelev_completion = as.Date(DATEPRELEV)) %>%
  group_by(PATID, EVTID) %>%
  mutate(
    .evt_start_date_completion = if (all(is.na(.date_prelev_completion))) as.Date(NA) else min(.date_prelev_completion, na.rm = TRUE),
    completion_evt_start_year = lubridate::year(.evt_start_date_completion)
  ) %>%
  ungroup() %>%
  mutate(
    nb_resultats_post = nb_resultats,
    completion_strategy = "raw",
    n_cells_filled = 0L,
    n_passes_touched = 0L
  ) %>%
  select(-.date_prelev_completion, -.evt_start_date_completion)

id_cols <- c(
  "PATID", "EVTID", "ELTID", "DATEPRELEV", "HEUREPRELEV", "souche_id",
  "naturepvt_norm", "bact_norm", "nb_resultats", "nb_resultats_post",
  "completion_strategy", "n_cells_filled", "n_passes_touched",
  "completion_evt_start_year"
)

candidate_atb <- if (exists("filtre_atb")) {
  intersect(filtre_atb, names(sir_wide_raw))
} else {
  setdiff(names(sir_wide_raw), id_cols)
}

is_atb_col <- function(x) {
  ux <- unique(na.omit(as.character(x)))
  length(ux) > 0 && all(ux %in% c("S", "R", "ZIT"))
}

atb_cols <- candidate_atb[vapply(sir_wide_raw[candidate_atb], is_atb_col, logical(1))]
stopifnot(length(atb_cols) > 0)

reorder_completion_dataset <- function(ds, atb_cols) {
  dataset_front_cols <- c(
    "PATID", "EVTID", "ELTID", "DATEPRELEV", "HEUREPRELEV",
    "souche_id", "naturepvt_norm", "bact_norm"
  )
  dataset_metric_cols <- c(
    "nb_resultats", "nb_resultats_post", "completion_strategy",
    "n_cells_filled", "n_passes_touched"
  )
  dataset_middle_cols <- setdiff(
    names(ds),
    c(dataset_front_cols, dataset_metric_cols, atb_cols)
  )
  dataset_order <- unique(c(
    dataset_front_cols,
    dataset_metric_cols,
    dataset_middle_cols,
    atb_cols
  ))
  ds %>% select(any_of(dataset_order))
}

sir_wide_raw <- reorder_completion_dataset(sir_wide_raw, atb_cols = atb_cols)

stopifnot(exists(".spares_sr_conflict_pair", mode = "function"))

zit_values <- "ZIT"
```

#### Load completion/helper contracts
Scripts are sourced once in `source_project_scripts` near the top of this notebook.
This section validates that required symbols are available and consistent.

```{r}
#| label: load_completion_helpers_and_checks
# Helper contract checks (documented in completion_helpers.R)
stopifnot(
  exists("is_sr", mode = "function"),
  exists("pair_gap_days_ok", mode = "function"),
  exists("transfer_sr", mode = "function"),
  exists("completeness_sort_key", mode = "function"),
  exists("complete_group_bidirectional", mode = "function"),
  exists("run_completion_strategy", mode = "function"),
  exists("build_raw_group_log", mode = "function"),
  exists("build_raw_row_log", mode = "function"),
  exists("validate_completion_payload", mode = "function"),
  exists("format_completion_contract_errors", mode = "function"),
  exists("normalize_strategy_name", mode = "function"),
  exists("compute_completion_fingerprint", mode = "function"),
  exists("build_completion_cache_meta", mode = "function"),
  exists("count_non_na_overwrites", mode = "function"),
  exists("count_non_na_to_na_losses", mode = "function"),
  exists("count_new_zit_from_fill", mode = "function"),
  exists("pct_rows_improved", mode = "function")
)

# Canonical conflict helper checks
stopifnot(
  .spares_sr_conflict_pair(
    c("S", NA, "ZIT", "R"),
    c("R", NA, "S", "R"),
    zit_values = zit_values
  ) == TRUE,
  .spares_sr_conflict_pair(
    c("S", NA),
    c(NA, "R"),
    zit_values = zit_values
  ) == FALSE,
  .spares_sr_conflict_pair(
    c("ZIT", "R"),
    c("S", NA),
    zit_values = zit_values
  ) == FALSE,
  .spares_sr_conflict_pair(
    c("I", "R"),
    c("S", "R"),
    zit_values = c("ZIT", "I")
  ) == FALSE
)
```

### Pre-completion gate

#### Structural sanity check
Question: Is the raw structure coherent enough to run completion safely at document, event, and patient levels?

Evidence: `ELTID`-level distribution/outlier tables plus `EVTID`/`PATID` structural distributions computed on `sir_wide_raw`.

```{r}
#| label: eltid_structural_audit
eltid_structural_base <- sir_wide_raw %>%
  group_by(ELTID) %>%
  summarise(
    n_rows = n(),
    n_naturepvt_norm = n_distinct(naturepvt_norm, na.rm = TRUE),
    n_bact_norm = n_distinct(bact_norm, na.rm = TRUE),
    n_date_prelev = n_distinct(DATEPRELEV, na.rm = TRUE),
    n_souche_id = n_distinct(souche_id, na.rm = TRUE),
    n_bact_souche_pairs = n_distinct(bact_norm, souche_id, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    across(
      c(n_naturepvt_norm, n_bact_norm, n_date_prelev, n_souche_id),
      ~ na_if(.x, 0L)
    )
  )

eltid_structural_distribution <- eltid_structural_base %>%
  count(
    n_rows, n_bact_norm, n_souche_id,
    name = "n_eltid", sort = TRUE
  ) %>%
  mutate(
    freq = 100 * n_eltid / sum(n_eltid),
    cumsum = cumsum(freq)
  )

eltid_structural_outliers <- eltid_structural_base %>%
  filter(
    n_date_prelev > 1 | n_bact_souche_pairs < n_rows
  ) %>%
  mutate(
    .date_anomaly_score = dplyr::coalesce(n_date_prelev, 0L),
    .pair_deficit = n_rows - n_bact_souche_pairs
  ) %>%
  arrange(desc(.date_anomaly_score), desc(.pair_deficit), desc(n_rows), ELTID) %>%
  select(-.date_anomaly_score, -.pair_deficit)

evtid_structural_base <- sir_wide_raw %>%
  group_by(EVTID) %>%
  summarise(
    n_rows = n(),
    n_eltid = n_distinct(ELTID, na.rm = TRUE),
    n_naturepvt_norm = n_distinct(naturepvt_norm, na.rm = TRUE),
    n_bact_norm = n_distinct(bact_norm, na.rm = TRUE),
    n_date_prelev = n_distinct(DATEPRELEV, na.rm = TRUE),
    n_souche_id = n_distinct(souche_id, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    across(
      c(n_eltid, n_naturepvt_norm, n_bact_norm, n_date_prelev, n_souche_id),
      ~ na_if(.x, 0L)
    )
  )

evtid_structural_distribution <- evtid_structural_base %>%
  count(
    n_rows, n_eltid, 
    # n_naturepvt_norm,
    n_bact_norm, n_date_prelev, n_souche_id,
    name = "n_evtid", sort = TRUE
  ) %>%
  mutate(
    freq = 100 * n_evtid / sum(n_evtid),
    cumsum = cumsum(freq)
  )

patid_structural_base <- sir_wide_raw %>%
  group_by(PATID,completion_evt_start_year) %>%
  summarise(
    n_rows = n(),
    n_evtid = n_distinct(EVTID, na.rm = TRUE),
    n_eltid = n_distinct(ELTID, na.rm = TRUE),
    n_naturepvt_norm = n_distinct(naturepvt_norm, na.rm = TRUE),
    n_bact_norm = n_distinct(bact_norm, na.rm = TRUE),
    n_date_prelev = n_distinct(DATEPRELEV, na.rm = TRUE),
    n_souche_id = n_distinct(souche_id, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    across(
      c(n_evtid, n_eltid, n_naturepvt_norm, n_bact_norm, n_date_prelev, n_souche_id),
      ~ na_if(.x, 0L)
    )
  )

patid_structural_distribution <- patid_structural_base %>% 
  group_by(completion_evt_start_year) %>% 
  count(
    n_rows, n_evtid, n_eltid, 
    # n_naturepvt_norm, 
    n_bact_norm, n_date_prelev, n_souche_id,
    name = "n_patid", sort = TRUE
  ) %>%
  mutate(
    freq = 100 * n_patid / sum(n_patid),
    cumsum = cumsum(freq)
  ) %>% 
    rename(year = completion_evt_start_year)

patid_structural_integrity <- patid_structural_distribution %>%
  group_by(year) %>%
  summarise(
    freq_sum = sum(freq),
    cumsum_max = max(cumsum),
    .groups = "drop"
  )

# Acceptance checks for audit table integrity.
stopifnot(
  sum(eltid_structural_distribution$n_eltid) == n_distinct(sir_wide_raw$ELTID),
  sum(eltid_structural_distribution$n_rows * eltid_structural_distribution$n_eltid) == nrow(sir_wide_raw),
  abs(sum(eltid_structural_distribution$freq) - 100) < 1e-8,
  abs(max(eltid_structural_distribution$cumsum) - 100) < 1e-8,
  sum(evtid_structural_distribution$n_evtid) == n_distinct(sir_wide_raw$EVTID),
  sum(evtid_structural_distribution$n_rows * evtid_structural_distribution$n_evtid) == nrow(sir_wide_raw),
  abs(sum(evtid_structural_distribution$freq) - 100) < 1e-8,
  abs(max(evtid_structural_distribution$cumsum) - 100) < 1e-8,
  # PATID structural table is built at PATID-year grain.
  sum(patid_structural_distribution$n_patid) == n_distinct(sir_wide_raw$PATID, sir_wide_raw$completion_evt_start_year),
  sum(patid_structural_distribution$n_rows * patid_structural_distribution$n_patid) == nrow(sir_wide_raw),
  all(abs(patid_structural_integrity$freq_sum - 100) < 1e-8),
  all(abs(patid_structural_integrity$cumsum_max - 100) < 1e-8),
  all(
    eltid_structural_outliers$n_date_prelev > 1 |
      eltid_structural_outliers$n_bact_souche_pairs < eltid_structural_outliers$n_rows
  )
)
```

```{r}
#| label: eltid_structural_distribution_table
num_cols <- names(eltid_structural_distribution)[
  vapply(eltid_structural_distribution, is.numeric, logical(1))
]

datatable_with_xlsx_button(
  data = eltid_structural_distribution,
  file_stem = "eltid_structural_distribution",
  caption = "ELTID structural distribution (document-level profile on sir_wide_raw)",
  rownames = FALSE,
  options = list(
    pageLength = 25,
    scrollX = TRUE,
    scrollY = "350px"
  ),
  format_round_cols = num_cols,
  digits = datatable_digits
)
```

Commentary: 
- ~91% of documents have only 1 row hence 1 souche
- ~99.5% of documents have only 1 row for each given bacteria
- only 200 documents (~0.5%) have multiple souches of the same bacteria



```{r}
#| label: evtid_structural_distribution_table

datatable_with_xlsx_button(
  data = evtid_structural_distribution,
  file_stem = "evtid_structural_distribution",
  caption = "EVTID structural distribution (event-level profile on sir_wide_raw)",
  rownames = FALSE,
  options = list(
    pageLength = 25,
    scrollX = TRUE,
    scrollY = "500px"
  ),
  format_round_cols = c("freq", "cumsum"),
  digits = datatable_digits
)
```

```{r}
evtid_linkage_ambiguity <- evtid_structural_distribution %>% 
  filter(n_bact_norm < n_rows, n_bact_norm == n_souche_id, n_rows > n_eltid, n_eltid > 1) %>%
  summarise(
    n_evtid_ambiguous = sum(n_evtid, na.rm = TRUE),
    pct_evtid_ambiguous = ifelse(
      n_distinct(sir_wide_raw$EVTID) > 0L,
      100 * n_evtid_ambiguous / n_distinct(sir_wide_raw$EVTID),
      NA_real_
    )
  )

evtid_linkage_ambiguity
```

Pattern used: `(n_bact_norm < n_rows, n_bact_norm == n_souche_id, n_rows > n_eltid, n_eltid > 1)`.
`r round(evtid_linkage_ambiguity$pct_evtid_ambiguous, 2)`% of EVTID groups show linkage ambiguity (`r evtid_linkage_ambiguity$n_evtid_ambiguous` / `r n_distinct(sir_wide_raw$EVTID)`).


```{r}
#| label: patid_structural_distribution_table
datatable_with_xlsx_button(
  data = patid_structural_distribution,
  file_stem = "patid_structural_distribution",
  caption = "PATID structural distribution (patient-level profile on sir_wide_raw)",
  filter = "top",
  rownames = FALSE,
  options = list(
    pageLength = 25,
    scrollX = TRUE,
    scrollY = "500px"
  ),
  format_round_cols = c("freq", "cumsum"),
  digits = datatable_digits
)
```


```{r}
# PATID with only 1 document
patid_structural_distribution %>% 
  filter(n_eltid==1) %>% 
  group_by(year) %>%
  summarise("PATID with only 1 document" = sum(n_patid)*100/n_distinct(sir_wide_raw$PATID))

# PATID 
patid_structural_distribution %>% 
  filter(n_bact_norm > n_souche_id, n_rows > n_eltid, n_eltid > 1) %>% 
  group_by(year) %>%
  summarise("PATID" = sum(n_patid)*100/n_distinct(sir_wide_raw$PATID))
```




```{r}
#| label: eltid_structural_outliers_table
datatable_with_xlsx_button(
  data = eltid_structural_outliers,
  file_stem = "eltid_structural_outliers",
  caption = "ELTID structural outliers (documents with date inconsistency or bact-souche deficits)",
  rownames = FALSE,
  options = list(
    pageLength = 25,
    scrollX = TRUE,
    scrollY = "500px"
  )
)
```



### Completion datasets (cache-first)

Cache note: this notebook uses explicit artifact caches (RDS files) for completion/dedup orchestration; this is independent from knitr chunk caching.

#### Strategy/spec setup

```{r}
#| label: completion_define_strategy_specs
completion_group_defs <- list(
  completion_date = c("PATID", "EVTID", "DATEPRELEV", "naturepvt_norm", "bact_norm"),
  completion_evtid = c("PATID", "EVTID", "naturepvt_norm", "bact_norm"),
  completion_patid_30d = c("PATID", "naturepvt_norm", "bact_norm"),
  completion_patid_year = c("PATID", "completion_evt_start_year", "naturepvt_norm", "bact_norm")
)

stopifnot(all(vapply(
  completion_group_defs,
  function(k) all(c("naturepvt_norm", "bact_norm") %in% k),
  logical(1)
)))

expected_dataset_prefix <- c(
  "PATID", "EVTID", "ELTID", "DATEPRELEV", "HEUREPRELEV", "souche_id", "naturepvt_norm", "bact_norm",
  "nb_resultats", "nb_resultats_post", "completion_strategy", "n_cells_filled", "n_passes_touched"
)
expected_row_cols <- c(
  ".row_id_global",
  "PATID", "EVTID", "ELTID", "DATEPRELEV", "HEUREPRELEV", "souche_id", "naturepvt_norm", "bact_norm",
  "nb_resultats_pre", "nb_resultats_post", "phenotype_class",
  "n_cells_filled", "n_passes_touched", "max_gap_used_days",
  "strategy", "max_days", "group_definition"
)
expected_group_cols <- c(
  "strategy", "max_days", "group_definition",
  "n_rows", "n_classes", "n_rows_filled", "n_cells_filled_total",
  "n_pairs_unique", "n_pairs_checked", "n_pairs_conflict", "n_iterations"
)
expected_dataset_names <- c("sir_wide_raw", names(completion_group_defs))
expected_log_names <- c("raw", names(completion_group_defs))

cache_candidates <- list(
  list(
    dataset = file.path("data", "completion_datasets"),
    logs = file.path("data", "completion_logs"),
    raw_row = file.path("data", "raw_row_log"),
    meta = file.path("data", "completion_cache_meta")
  )
)

completion_cache_source <- NA_character_
completion_datasets <- NULL
completion_logs <- NULL
raw_row_log <- NULL
completion_cache_meta <- NULL
completion_cache_decision <- NA_character_
```

#### Cache control

```{r}
#| label: completion_cache_control
recompute_completion <- orchidee_config$cache$recompute_completion
```

#### Cache resolution

```{r}
#| label: completion_cache_load

completion_cache_meta_current <- build_completion_cache_meta(sir_wide_raw, atb_cols)

if (!isTRUE(recompute_completion)) {
  for (candidate in cache_candidates) {
    cache_paths <- unlist(candidate, use.names = FALSE)
    if (!all(file.exists(cache_paths))) {
      next
    }

    loaded <- tryCatch(
      list(
        completion_datasets = readRDS(candidate$dataset),
        completion_logs = readRDS(candidate$logs),
        raw_row_log = readRDS(candidate$raw_row),
        completion_cache_meta = readRDS(candidate$meta)
      ),
      error = function(e) NULL
    )

    if (is.null(loaded)) {
      next
    }
    if (!all(expected_dataset_names %in% names(loaded$completion_datasets))) {
      next
    }
    if (!all(expected_log_names %in% names(loaded$completion_logs))) {
      next
    }
    if (!all(atb_cols %in% names(loaded$completion_datasets$sir_wide_raw))) {
      next
    }
    if (nrow(loaded$completion_datasets$sir_wide_raw) != nrow(sir_wide_raw)) {
      next
    }
    if (n_distinct(loaded$completion_datasets$sir_wide_raw$ELTID) != n_distinct(sir_wide_raw$ELTID)) {
      next
    }
    if (!is.list(loaded$completion_cache_meta)) {
      next
    }
    required_completion_meta_fields <- c(
      "fingerprint", "atb_cols", "n_rows", "n_distinct_eltid",
      "phenotype_cols", "script_hashes"
    )
    if (!all(required_completion_meta_fields %in% names(loaded$completion_cache_meta))) {
      next
    }
    if (!identical(as.character(loaded$completion_cache_meta$fingerprint), completion_cache_meta_current$fingerprint)) {
      next
    }
    if (!identical(loaded$completion_cache_meta$atb_cols, atb_cols)) {
      next
    }
    if (!identical(sort(as.character(loaded$completion_cache_meta$phenotype_cols)), sort(as.character(completion_cache_meta_current$phenotype_cols)))) {
      next
    }
    current_completion_hashes <- unlist(completion_cache_meta_current$script_hashes, use.names = TRUE)
    loaded_completion_hashes <- unlist(loaded$completion_cache_meta$script_hashes, use.names = TRUE)
    if (!identical(loaded_completion_hashes[names(current_completion_hashes)], current_completion_hashes)) {
      next
    }
    if (!isTRUE(loaded$completion_cache_meta$n_rows == nrow(sir_wide_raw))) {
      next
    }
    if (!isTRUE(loaded$completion_cache_meta$n_distinct_eltid == n_distinct(sir_wide_raw$ELTID))) {
      next
    }

    loaded_contract <- validate_completion_payload(
      completion_datasets = loaded$completion_datasets,
      completion_logs = loaded$completion_logs,
      atb_cols = atb_cols,
      expected_dataset_prefix = expected_dataset_prefix,
      expected_row_cols = expected_row_cols,
      expected_group_cols = expected_group_cols
    )
    if (!loaded_contract$ok) {
      next
    }
    expected_group_definitions <- vapply(
      completion_group_defs,
      function(keys) paste(keys, collapse = ", "),
      character(1)
    )
    strategy_contract_ok <- all(vapply(names(expected_group_definitions), function(nm) {
      entry <- loaded$completion_logs[[nm]]
      if (is.null(entry) || is.null(entry$group_log) || is.null(entry$row_log)) {
        return(FALSE)
      }
      expected_def <- expected_group_definitions[[nm]]
      group_ok <- if (nrow(entry$group_log) == 0L) {
        TRUE
      } else {
        all(entry$group_log$group_definition == expected_def)
      }
      row_ok <- if (nrow(entry$row_log) == 0L) {
        TRUE
      } else {
        all(entry$row_log$group_definition == expected_def)
      }
      group_ok && row_ok
    }, logical(1)))
    if (!strategy_contract_ok) {
      next
    }

    completion_datasets <- loaded$completion_datasets
    completion_logs <- loaded$completion_logs
    raw_row_log <- loaded$raw_row_log
    completion_cache_meta <- loaded$completion_cache_meta
    completion_cache_source <- candidate$dataset
    break
  }
} else {
  completion_cache_source <- "forced_recompute"
}
```

#### Recompute fallback

```{r}
#| label: completion_recompute_if_needed

if (isTRUE(recompute_completion)) {
  completion_datasets <- NULL
  completion_logs <- NULL
  raw_row_log <- NULL
  completion_cache_meta <- NULL
  completion_cache_source <- "forced_recompute"
  cat("completion cache bypass: recompute_completion=TRUE\n")
}

if (!isTRUE(recompute_completion) && !is.null(completion_datasets) && !is.null(completion_logs) && !is.null(raw_row_log)) {
  message("Loaded completion artifacts from ", completion_cache_source)
  completion_cache_decision <- "loaded"
} else {
  if (isTRUE(recompute_completion)) {
    message("Force recompute requested; computing completion datasets.")
  } else {
    message("No valid completion cache found; computing completion datasets.")
  }
  if (!"completion_evt_start_year" %in% names(sir_wide_raw)) {
    stopifnot(all(c("PATID", "EVTID", "DATEPRELEV") %in% names(sir_wide_raw)))
    sir_wide_raw <- sir_wide_raw %>%
      mutate(.date_prelev_completion = as.Date(DATEPRELEV)) %>%
      group_by(PATID, EVTID) %>%
      mutate(
        .evt_start_date_completion = if (all(is.na(.date_prelev_completion))) as.Date(NA) else min(.date_prelev_completion, na.rm = TRUE),
        completion_evt_start_year = lubridate::year(.evt_start_date_completion)
      ) %>%
      ungroup() %>%
      select(-.date_prelev_completion, -.evt_start_date_completion)
  }
  stopifnot(exists("reorder_completion_dataset", mode = "function"))
  sir_wide_raw <- reorder_completion_dataset(sir_wide_raw, atb_cols = atb_cols)

  res_completion_date <- run_completion_strategy(
    df = sir_wide_raw,
    group_keys = completion_group_defs$completion_date,
    atb_cols = atb_cols,
    strategy_name = "completion_date",
    max_days = Inf,
    zit_values = zit_values
  )

  res_completion_evtid <- run_completion_strategy(
    df = sir_wide_raw,
    group_keys = completion_group_defs$completion_evtid,
    atb_cols = atb_cols,
    strategy_name = "completion_evtid",
    max_days = Inf,
    zit_values = zit_values
  )

  res_completion_patid_30d <- run_completion_strategy(
    df = sir_wide_raw,
    group_keys = completion_group_defs$completion_patid_30d,
    atb_cols = atb_cols,
    strategy_name = "completion_patid_30d",
    max_days = 30,
    zit_values = zit_values
  )

  res_completion_patid_year <- run_completion_strategy(
    df = sir_wide_raw,
    group_keys = completion_group_defs$completion_patid_year,
    atb_cols = atb_cols,
    strategy_name = "completion_patid_year",
    max_days = Inf,
    zit_values = zit_values
  )

  raw_group_log <- build_raw_group_log(sir_wide_raw)
  raw_row_log <- build_raw_row_log(sir_wide_raw)

  completion_datasets <- list(
    sir_wide_raw = sir_wide_raw,
    completion_date = res_completion_date$dataset,
    completion_evtid = res_completion_evtid$dataset,
    completion_patid_30d = res_completion_patid_30d$dataset,
    completion_patid_year = res_completion_patid_year$dataset
  )

  completion_logs <- list(
    raw = list(group_log = raw_group_log, row_log = raw_row_log),
    completion_date = list(group_log = res_completion_date$group_log, row_log = res_completion_date$row_log),
    completion_evtid = list(group_log = res_completion_evtid$group_log, row_log = res_completion_evtid$row_log),
    completion_patid_30d = list(group_log = res_completion_patid_30d$group_log, row_log = res_completion_patid_30d$row_log),
    completion_patid_year = list(group_log = res_completion_patid_year$group_log, row_log = res_completion_patid_year$row_log)

  )
  completion_cache_meta <- build_completion_cache_meta(sir_wide_raw, atb_cols)

  saveRDS(completion_datasets, file.path("data", "completion_datasets"))
  saveRDS(completion_logs, file.path("data", "completion_logs"))
  saveRDS(raw_row_log, file.path("data", "raw_row_log"))
  saveRDS(completion_cache_meta, file.path("data", "completion_cache_meta"))
  completion_cache_source <- "recomputed"
  completion_cache_decision <- "recomputed"
  message("Saved completion artifacts to: data/completion_datasets, data/completion_logs, data/raw_row_log, data/completion_cache_meta")
}

stopifnot(
  !is.null(completion_datasets),
  !is.null(completion_logs),
  !is.null(raw_row_log)
)
cat("recompute_completion: ", isTRUE(recompute_completion), "\n", sep = "")
cat("completion_cache_source: ", completion_cache_source, "\n", sep = "")
cat("completion_cache_decision: ", completion_cache_decision, "\n", sep = "")
cat("Completion datasets ready: ", paste(names(completion_datasets), collapse = ", "), "\n", sep = "")
```

#### Payload validation

```{r}
#| label: completion_payload_validation

completion_contract <- validate_completion_payload(
  completion_datasets = completion_datasets,
  completion_logs = completion_logs,
  atb_cols = atb_cols,
  expected_dataset_prefix = expected_dataset_prefix,
  expected_row_cols = expected_row_cols,
  expected_group_cols = expected_group_cols
)

if (!completion_contract$ok) {
  stop(paste(format_completion_contract_errors(completion_contract), collapse = "\n"), call. = FALSE)
}

names(completion_datasets)
```

## Completion Audits And Decisions

### Completion safety audit

Question: After completion, do we preserve structural and value-safety invariants?

Evidence: schema contract checks above + cross-dataset safety metrics below.

```{r}
#| label: completion_safety_checks
completion_checks <- purrr::imap_dfr(completion_datasets, function(ds, nm) {
  tibble(
    dataset = nm,
    n_rows = nrow(ds),
    n_eltid = n_distinct(ds$ELTID),
    # Structural invariants:
    same_n_rows_as_raw = nrow(ds) == nrow(sir_wide_raw),
    same_n_eltid_as_raw = n_distinct(ds$ELTID) == n_distinct(sir_wide_raw$ELTID),
    # Value-safety invariants:
    non_na_overwrite_count = count_non_na_overwrites(sir_wide_raw, ds, atb_cols),
    non_na_to_na_loss_count = count_non_na_to_na_losses(sir_wide_raw, ds, atb_cols),
    filled_with_zit_count = count_new_zit_from_fill(sir_wide_raw, ds, atb_cols),
    # Completion monotonicity:
    monotonic_nb_resultats = all(ds$nb_resultats_post >= ds$nb_resultats),
    # Interpretability KPI:
    # pct_rows_improved = 100 * mean(nb_resultats_post > nb_resultats).
    pct_rows_improved = pct_rows_improved(ds)
  )
})

patid_30d_gap_ok <- completion_logs$completion_patid_30d$row_log %>%
  filter(n_cells_filled > 0) %>%
  summarise(ok = all(is.na(max_gap_used_days) | max_gap_used_days <= 30)) %>%
  pull(ok)

if (length(patid_30d_gap_ok) == 0) patid_30d_gap_ok <- TRUE

completion_safety_status <- tibble::tribble(
  ~check, ~ok,
  "same_n_rows_as_raw", all(completion_checks$same_n_rows_as_raw),
  "same_n_eltid_as_raw", all(completion_checks$same_n_eltid_as_raw),
  "non_na_overwrite_count_zero", all(completion_checks$non_na_overwrite_count == 0),
  "non_na_to_na_loss_count_zero", all(completion_checks$non_na_to_na_loss_count == 0),
  "filled_with_zit_count_zero", all(completion_checks$filled_with_zit_count == 0),
  "monotonic_nb_resultats", all(completion_checks$monotonic_nb_resultats),
  "patid_30d_gap_ok", isTRUE(patid_30d_gap_ok)
)

completion_safety_ok <- all(completion_safety_status$ok)
stopifnot(completion_safety_ok)

show_audit_table(
  data = completion_checks,
  file_stem = "completion_safety_checks",
  caption = "Completion safety checks",
  page_length = 10,
  scroll_y = "320px",
  digits = datatable_digits
)
```

Decision: completion safety is accepted only when `completion_safety_ok` is `TRUE`.\

Comments: only a small number of rows seems to benefit from cross-row completion within a class. The larger the window the bigger this percentage.

```{r}
#| label: conflict_consistency_compute
check_conflict_consistency <- function(dataset, dataset_name, atb_cols, zit_values = "ZIT", max_rows = 120L, seed = 42L) {
  stopifnot(all(atb_cols %in% names(dataset)))

  n <- nrow(dataset)
  if (n <= 1L) {
    return(tibble(
      dataset = dataset_name,
      n_rows_sampled = n,
      n_pairs_tested = 0L,
      n_mismatch = 0L
    ))
  }

  set.seed(seed)
  idx <- if (n > max_rows) sample.int(n, max_rows) else seq_len(n)
  g <- dataset[idx, , drop = FALSE]

  D <- .spares_discord_matrix(g, atb_cols = atb_cols, zit_values = zit_values)
  pair_idx <- which(upper.tri(D), arr.ind = TRUE)

  if (nrow(pair_idx) == 0L) {
    return(tibble(
      dataset = dataset_name,
      n_rows_sampled = nrow(g),
      n_pairs_tested = 0L,
      n_mismatch = 0L
    ))
  }

  mismatch <- 0L
  for (k in seq_len(nrow(pair_idx))) {
    i <- pair_idx[k, 1]
    j <- pair_idx[k, 2]

    a <- as.character(unlist(g[i, atb_cols, drop = FALSE], use.names = FALSE))
    b <- as.character(unlist(g[j, atb_cols, drop = FALSE], use.names = FALSE))

    pair_flag <- .spares_sr_conflict_pair(a, b, zit_values = zit_values)
    matrix_flag <- isTRUE(D[i, j])

    if (!identical(pair_flag, matrix_flag)) mismatch <- mismatch + 1L
  }

  tibble(
    dataset = dataset_name,
    n_rows_sampled = nrow(g),
    n_pairs_tested = nrow(pair_idx),
    n_mismatch = mismatch
  )
}

# Canonical single-source conflict consistency object used by diagnostics below.
conflict_consistency <- purrr::imap_dfr(
  completion_datasets,
  ~ check_conflict_consistency(.x, .y, atb_cols = atb_cols, zit_values = zit_values)
)

stopifnot(all(conflict_consistency$n_mismatch == 0L))
```

```{r}
#| label: completion_diagnostics
# Aggregate per-group completion metrics to one line per dataset.
# Sums are additive counts; mean_iterations is averaged over groups.
group_kpis <- imap_dfr(completion_logs, ~{
  .x$group_log %>%
    summarise(
      n_groups = n(),
      n_rows = sum(n_rows, na.rm = TRUE),
      n_rows_filled = sum(n_rows_filled, na.rm = TRUE),
      n_cells_filled_total = sum(n_cells_filled_total, na.rm = TRUE),
      n_pairs_checked = sum(n_pairs_checked, na.rm = TRUE),
      n_pairs_conflict = sum(n_pairs_conflict, na.rm = TRUE),
      mean_iterations = mean(n_iterations, na.rm = TRUE)
    ) %>%
    mutate(dataset = normalize_strategy_name(.y))
}) %>% select(dataset, everything())

# Inspect highest-impact rows in trusted mapped sample types.
top_fills <- imap_dfr(completion_logs, ~{
  .x$row_log %>% mutate(dataset = normalize_strategy_name(.y))
}) %>%
  filter(n_cells_filled > 0, naturepvt_norm %in% c("hemoculture","urines")) %>%
  arrange(desc(n_cells_filled), desc(max_gap_used_days)) %>%
  select(
    dataset, PATID, EVTID, ELTID, naturepvt_norm, bact_norm,
    nb_resultats_pre, nb_resultats_post,
    n_cells_filled, n_passes_touched, max_gap_used_days
  ) %>%
  head(30)

# Canonical per-strategy impact synthesis; completion_gate is a compact projection of this table.
row_drift_vs_raw <- imap_dfr(completion_logs, function(log_obj, nm) {
  dataset_nm <- normalize_strategy_name(nm)
  n_rows_log <- nrow(log_obj$row_log)
  n_diff <- sum(log_obj$row_log$n_cells_filled > 0L, na.rm = TRUE)
  tibble(
    dataset = dataset_nm,
    n_diff_vs_raw = as.integer(n_diff),
    pct_diff_vs_raw = if (n_rows_log > 0L) 100 * n_diff / n_rows_log else NA_real_
  )
})

strategy_impact_summary <- completion_checks %>%
  left_join(
    group_kpis %>% select(dataset, n_rows_filled, n_cells_filled_total, n_pairs_conflict, mean_iterations),
    by = "dataset"
  ) %>%
  left_join(row_drift_vs_raw, by = "dataset") %>%
  arrange(match(dataset, names(completion_datasets)))

# Compact QA gate shown in the report header.
completion_gate <- strategy_impact_summary %>%
  select(
    dataset, n_rows, n_eltid, pct_rows_improved,
    non_na_overwrite_count, non_na_to_na_loss_count, filled_with_zit_count, monotonic_nb_resultats
  )

stopifnot(exists("conflict_consistency"))

completion_diagnostics <- list(
  completion_gate = completion_gate,
  # Compatibility alias: keep this slot name stable for downstream snippets.
  conflict_consistency = conflict_consistency,
  group_kpis = group_kpis,
  top_fills = top_fills
)

show_audit_table(
  data = completion_diagnostics$completion_gate,
  file_stem = "completion_gate",
  caption = "Completion gate",
  page_length = 10,
  scroll_y = "260px"
)

show_audit_table(
  data = completion_diagnostics$conflict_consistency,
  file_stem = "completion_conflict_consistency",
  caption = "Conflict consistency",
  page_length = 10,
  scroll_y = "260px"
)

show_audit_table(
  data = completion_diagnostics$group_kpis,
  file_stem = "completion_group_kpis",
  caption = "Completion group KPIs",
  page_length = 10,
  scroll_y = "260px"
)

show_audit_table(
  data = completion_diagnostics$top_fills,
  file_stem = "completion_top_fills",
  caption = "Top fills (hemoculture/urines)",
  page_length = 30,
  scroll_y = "420px"
)

invisible(completion_diagnostics)
```

### Strategy impact audit

Question: after safety is validated, does completion materially change rows and how much by strategy?

Evidence: `strategy_impact_summary` merges safety KPIs, group-level fill activity, and row-level drift vs raw.

```{r}
#| label: strategy_impact_audit
stopifnot(exists("strategy_impact_summary"))

show_audit_table(
  data = strategy_impact_summary,
  file_stem = "strategy_impact_summary",
  caption = "Strategy impact summary",
  page_length = 10,
  scroll_y = "280px"
)
```

Decision: completion strategy is considered useful when it keeps safety invariants and shows non-trivial `n_diff_vs_raw` plus `pct_rows_improved`.

### Strategy drift audit

Question: do completion strategies produce equivalent row content or meaningful drift?

Evidence: `strategy_drift_summary` compares row equality between strategy pairs across all common columns except `completion_strategy`.

```{r}
#| label: strategy_drift_audit
compute_strategy_drift <- function(a, b, label_a, label_b, compare_cols = NULL) {
  if (is.null(compare_cols)) {
    compare_cols <- setdiff(intersect(names(a), names(b)), "completion_strategy")
  }

  x <- a %>% select(any_of(compare_cols))
  y <- b %>% select(any_of(compare_cols))

  same_rows <- semi_join(x, y, by = compare_cols)
  diff_rows <- anti_join(x, y, by = compare_cols)

  n_total <- nrow(x)
  n_same <- nrow(same_rows)
  n_diff <- nrow(diff_rows)

  tibble(
    comparison = paste(label_a, "->", label_b),
    n_total = n_total,
    n_same = n_same,
    n_diff = n_diff,
    pct_same = if (n_total > 0L) 100 * n_same / n_total else NA_real_,
    pct_diff = if (n_total > 0L) 100 * n_diff / n_total else NA_real_
  )
}

strategy_names <- setdiff(names(completion_datasets), "sir_wide_raw")
strategy_pairs <- if (length(strategy_names) >= 2L) {
  as.data.frame(t(combn(strategy_names, 2L)), stringsAsFactors = FALSE) %>%
    setNames(c("from", "to")) %>%
    as_tibble()
} else {
  tibble(from = character(), to = character())
}

strategy_drift_summary <- purrr::pmap_dfr(
  strategy_pairs,
  function(from, to) {
    compute_strategy_drift(
      a = completion_datasets[[from]],
      b = completion_datasets[[to]],
      label_a = from,
      label_b = to
    )
  }
)

show_audit_table(
  data = strategy_drift_summary,
  file_stem = "strategy_drift_summary",
  caption = "Strategy drift summary",
  page_length = 20,
  scroll_y = "360px",
  format_round_cols = c("pct_same", "pct_diff"),
  digits = datatable_digits
)
```

Decision: strategy choice matters when `strategy_drift_summary$n_diff > 0` for any comparison.

### Conflict consistency audit
Question: are both conflict detectors internally consistent across sampled pairs?

Evidence: `conflict_consistency` reports sampled row counts, tested pairs, and mismatches between helper and matrix implementations.

```{r}
#| label: conflict_consistency_audit
stopifnot(exists("conflict_consistency"))
stopifnot(all(conflict_consistency$n_mismatch == 0L))

show_audit_table(
  data = conflict_consistency,
  file_stem = "conflict_consistency",
  caption = "Conflict consistency audit",
  page_length = 10,
  scroll_y = "280px"
)
```

Decision: conflict implementation is accepted only if `n_mismatch == 0` for every dataset.

# Dedup

## Execution

Question: how do we execute SPARES-aligned dedup across datasets on the sample-level microbiology TA/DE perimeter?

Evidence: `dedup_results` and `dedup_summary` materialize patient-year scoped dedup outputs and core invariants.

Cache note: dedup execution below is cache-first via `dedup_results`/`dedup_cache_meta`, independent from knitr chunk cache.

```{r}
#| label: dedup_cache_setup
dedup_scope_defs <- list(
  by_type = c("PATID", "dedup_year", "naturepvt_norm", "bact_norm"),
  global = c("PATID", "dedup_year", "bact_norm")
)

prepare_dedup_dataset <- function(ds) {
  stopifnot(is.data.frame(ds), all(c("PATID", "DATEPRELEV") %in% names(ds)))
  out <- ds %>%
    mutate(dedup_year = lubridate::year(as.Date(DATEPRELEV)))
  if (any(is.na(out$dedup_year))) {
    stop(
      "Missing dedup_year detected: DATEPRELEV must be non-missing/date-coercible for PATID+year dedup.",
      call. = FALSE
    )
  }
  out
}

completion_datasets_dedup <- purrr::map(completion_datasets, prepare_dedup_dataset)

stopifnot(
  all(c("PATID", "dedup_year", "naturepvt_norm", "bact_norm") %in% dedup_scope_defs$by_type),
  all(c("PATID", "dedup_year", "bact_norm") %in% dedup_scope_defs$global),
  !"naturepvt_norm" %in% dedup_scope_defs$global
)

dedup_cache_candidates <- list(
  list(
    results = file.path("data", "dedup_results"),
    meta = file.path("data", "dedup_cache_meta")
  )
)

dedup_cache_source <- NA_character_
dedup_results <- NULL
dedup_cache_meta <- NULL
dedup_cache_decision <- NA_character_

build_dedup_cache_meta <- function(completion_datasets_dedup,
                                   dedup_scope_defs,
                                   atb_cols,
                                   zit_values,
                                   completion_cache_meta) {
  stopifnot(is.list(completion_cache_meta), !is.null(completion_cache_meta$fingerprint))

  dataset_signatures <- purrr::imap_dfr(
    completion_datasets_dedup,
    function(ds, nm) {
      tibble(
        dataset = nm,
        n_rows = nrow(ds),
        n_distinct_eltid = dplyr::n_distinct(ds$ELTID)
      )
    }
  ) %>%
    arrange(dataset)

  script_paths <- c(
    spares_dedup = orchidee_resolve_script_path(
      "spares_dedup.R",
      "script required for dedup cache fingerprint"
    ),
    spares_shared_primitives = orchidee_resolve_script_path(
      "spares_shared_primitives.R",
      "script required for dedup cache fingerprint"
    ),
    phenotype_flag_helpers = orchidee_resolve_script_path(
      "phenotype_flag_helpers.R",
      "script required for dedup cache fingerprint"
    )
  )
  script_hashes <- as.list(as.character(unname(tools::md5sum(script_paths))))
  names(script_hashes) <- names(script_paths)

  payload <- list(
    completion_fingerprint = as.character(completion_cache_meta$fingerprint),
    atb_cols = atb_cols,
    zit_values = zit_values,
    dedup_scope_defs = dedup_scope_defs,
    dataset_signatures = dataset_signatures,
    script_hashes = script_hashes
  )

  tmp_file <- tempfile(pattern = "dedup_fingerprint_", fileext = ".rds")
  on.exit(unlink(tmp_file), add = TRUE)
  saveRDS(payload, tmp_file, version = 2)
  fingerprint <- as.character(unname(tools::md5sum(tmp_file)))

  list(
    fingerprint = fingerprint,
    completion_fingerprint = as.character(completion_cache_meta$fingerprint),
    atb_cols = atb_cols,
    zit_values = zit_values,
    dedup_scope_defs = dedup_scope_defs,
    dataset_signatures = dataset_signatures,
    script_hashes = script_hashes,
    created_at = format(Sys.time(), tz = "UTC", usetz = TRUE)
  )
}

dedup_cache_meta_current <- build_dedup_cache_meta(
  completion_datasets_dedup = completion_datasets_dedup,
  dedup_scope_defs = dedup_scope_defs,
  atb_cols = atb_cols,
  zit_values = zit_values,
  completion_cache_meta = completion_cache_meta
)

expected_dedup_dataset_names <- names(completion_datasets_dedup)
expected_dedup_scope_names <- names(dedup_scope_defs)
```

```{r}
#| label: dedup_cache_control
recompute_dedup <- orchidee_config$cache$recompute_dedup
```

```{r}
#| label: dedup_cache_load
if (!isTRUE(recompute_dedup)) {
  for (candidate in dedup_cache_candidates) {
    cache_paths <- unlist(candidate, use.names = FALSE)
    if (!all(file.exists(cache_paths))) {
      next
    }

    loaded <- tryCatch(
      list(
        dedup_results = readRDS(candidate$results),
        dedup_cache_meta = readRDS(candidate$meta)
      ),
      error = function(e) NULL
    )

    if (is.null(loaded) || !is.list(loaded$dedup_results) || !is.list(loaded$dedup_cache_meta)) {
      next
    }
    if (!all(expected_dedup_dataset_names %in% names(loaded$dedup_results))) {
      next
    }

    structure_ok <- all(vapply(expected_dedup_dataset_names, function(dataset_name) {
      scope_list <- loaded$dedup_results[[dataset_name]]
      if (is.null(scope_list) || !is.list(scope_list) || !all(expected_dedup_scope_names %in% names(scope_list))) {
        return(FALSE)
      }
      all(vapply(expected_dedup_scope_names, function(scope_name) {
        res <- scope_list[[scope_name]]
        if (is.null(res) || !is.list(res) || !all(c("dedup", "class_map", "episode_summary", "audit") %in% names(res))) {
          return(FALSE)
        }
        if (!is.data.frame(res$dedup) || !is.data.frame(res$class_map) || !is.data.frame(res$episode_summary) || !is.data.frame(res$audit)) {
          return(FALSE)
        }
        scope_keys <- dedup_scope_defs[[scope_name]]
        dedup_ok <- all(c(scope_keys, "phenotype_class") %in% names(res$dedup))
        class_ok <- all(c(scope_keys, "phenotype_class") %in% names(res$class_map))
        episode_ok <- all(c(scope_keys, "n_docs", "n_classes", "has_multiple_classes") %in% names(res$episode_summary))
        audit_ok <- all(c(scope_keys, "order_sensitive", "n_within_class_discord_pairs") %in% names(res$audit))
        dedup_ok && class_ok && episode_ok && audit_ok
      }, logical(1)))
    }, logical(1)))
    if (!structure_ok) {
      next
    }

    required_meta_fields <- c(
      "fingerprint", "completion_fingerprint", "atb_cols", "zit_values",
      "dedup_scope_defs", "dataset_signatures", "script_hashes"
    )
    if (!all(required_meta_fields %in% names(loaded$dedup_cache_meta))) {
      next
    }

    if (!identical(as.character(loaded$dedup_cache_meta$fingerprint), dedup_cache_meta_current$fingerprint)) {
      next
    }
    if (!identical(as.character(loaded$dedup_cache_meta$completion_fingerprint), dedup_cache_meta_current$completion_fingerprint)) {
      next
    }
    if (!identical(loaded$dedup_cache_meta$atb_cols, atb_cols)) {
      next
    }
    if (!identical(loaded$dedup_cache_meta$zit_values, zit_values)) {
      next
    }
    if (!identical(loaded$dedup_cache_meta$dedup_scope_defs, dedup_scope_defs)) {
      next
    }

    current_sig <- dedup_cache_meta_current$dataset_signatures %>%
      transmute(dataset = as.character(dataset), n_rows = as.integer(n_rows), n_distinct_eltid = as.integer(n_distinct_eltid)) %>%
      arrange(dataset)
    loaded_sig <- loaded$dedup_cache_meta$dataset_signatures
    if (!is.data.frame(loaded_sig) || !all(c("dataset", "n_rows", "n_distinct_eltid") %in% names(loaded_sig))) {
      next
    }
    loaded_sig <- loaded_sig %>%
      transmute(dataset = as.character(dataset), n_rows = as.integer(n_rows), n_distinct_eltid = as.integer(n_distinct_eltid)) %>%
      arrange(dataset)
    if (!identical(loaded_sig, current_sig)) {
      next
    }

    current_script_hashes <- unlist(dedup_cache_meta_current$script_hashes, use.names = TRUE)
    loaded_script_hashes <- unlist(loaded$dedup_cache_meta$script_hashes, use.names = TRUE)
    if (!identical(loaded_script_hashes[names(current_script_hashes)], current_script_hashes)) {
      next
    }

    dedup_results <- loaded$dedup_results
    dedup_cache_meta <- loaded$dedup_cache_meta
    dedup_cache_source <- candidate$results
    break
  }
} else {
  dedup_cache_source <- "forced_recompute"
}
```

```{r}
#| label: dedup_execution
if (isTRUE(recompute_dedup)) {
  dedup_results <- NULL
  dedup_cache_meta <- NULL
  dedup_cache_source <- "forced_recompute"
  cat("dedup cache bypass: recompute_dedup=TRUE\n")
}

if (!isTRUE(recompute_dedup) && !is.null(dedup_results)) {
  message("Loaded dedup artifacts from ", dedup_cache_source)
  dedup_cache_decision <- "loaded"
} else {
  if (isTRUE(recompute_dedup)) {
    message("Force recompute requested; computing dedup outputs.")
  } else {
    message("No valid dedup cache found; computing dedup outputs.")
  }
  dedup_results <- purrr::imap(
    completion_datasets_dedup,
    function(ds, dataset_name) {
      representative_completeness_col <- if (identical(dataset_name, "sir_wide_raw")) {
        "nb_resultats"
      } else {
        "nb_resultats_post"
      }
      purrr::imap(
        dedup_scope_defs,
        function(scope_keys, scope_name) {
          stopifnot(all(scope_keys %in% names(ds)))
          spares_dedup(
            df = ds,
            atb_cols = atb_cols,
            group_keys = scope_keys,
            time_col = "HEUREPRELEV",
            date_col = "DATEPRELEV",
            document_id_col = "ELTID",
            completeness_col = representative_completeness_col,
            zit_values = zit_values,
            keep_class_members = TRUE,
            keep_audit = TRUE
          )
        }
      )
    }
  )
  dedup_cache_meta <- dedup_cache_meta_current
  saveRDS(dedup_results, file.path("data", "dedup_results"))
  saveRDS(dedup_cache_meta, file.path("data", "dedup_cache_meta"))
  dedup_cache_source <- "recomputed"
  dedup_cache_decision <- "recomputed"
  message("Saved dedup artifacts to: data/dedup_results, data/dedup_cache_meta")
}

stopifnot(!is.null(dedup_results))
cat("recompute_dedup: ", isTRUE(recompute_dedup), "\n", sep = "")
cat("dedup_cache_source: ", dedup_cache_source, "\n", sep = "")
cat("dedup_cache_decision: ", dedup_cache_decision, "\n", sep = "")

# One-row summary per (dataset, scope):
# - n_input_rows: input size before dedup.
# - n_retained_rows: representatives kept after dedup.
# - pct_episodes_multi_class: 100 * mean(has_multiple_classes).
# - order_sensitive_rate_pct: 100 * mean(order_sensitive).
# - n_within_class_discord_pairs must stay 0.
dedup_summary <- purrr::imap_dfr(
  dedup_results,
  function(scope_list, dataset_name) {
    purrr::imap_dfr(
      scope_list,
      function(res, scope_name) {
        episode <- res$episode_summary
        audit <- res$audit
        tibble(
          dataset = dataset_name,
          scope = scope_name,
          n_input_rows = nrow(completion_datasets_dedup[[dataset_name]]),
          n_retained_rows = nrow(res$dedup),
          n_episodes = nrow(episode),
          pct_episodes_multi_class = if (nrow(episode) > 0L) mean(episode$has_multiple_classes, na.rm = TRUE) * 100 else NA_real_,
          order_sensitive_rate_pct = if (!is.null(audit) && nrow(audit) > 0L) mean(audit$order_sensitive, na.rm = TRUE) * 100 else NA_real_,
          n_within_class_discord_pairs = if (!is.null(audit) && nrow(audit) > 0L) sum(audit$n_within_class_discord_pairs, na.rm = TRUE) else 0L
        )
      }
    )
  }
)

scope_sanity <- dedup_summary %>%
  select(dataset, scope, n_retained_rows) %>%
  pivot_wider(names_from = scope, values_from = n_retained_rows)

stopifnot(all(scope_sanity$global <= scope_sanity$by_type))
stopifnot(all(dedup_summary$n_within_class_discord_pairs == 0L))

show_audit_table(
  data = dedup_summary,
  file_stem = "dedup_summary",
  caption = "Dedup summary",
  page_length = 10,
  scroll_y = "320px",
  digits = datatable_digits
)
```

## Dedup Audits And Decisions

### Dedup impact vs raw

Question: after dedup execution, how much do dedup outcomes shift relative to raw baseline?

Evidence: `dedup_impact_summary` compares retained rows, key audit rates, and representative identity drift by scope.

```{r}
#| label: dedup_impact_audit

rep_key <- c(
  "PATID", "EVTID", "ELTID", "souche_id",
  "naturepvt_norm", "bact_norm", "DATEPRELEV", "HEUREPRELEV"
)

compute_rep_identity_drift <- function(res_raw, res_new, rep_key) {
  stopifnot(all(rep_key %in% names(res_raw$dedup)), all(rep_key %in% names(res_new$dedup)))
  a <- res_raw$dedup %>% distinct(across(all_of(rep_key)))
  b <- res_new$dedup %>% distinct(across(all_of(rep_key)))
  removed <- anti_join(a, b, by = rep_key)
  added <- anti_join(b, a, by = rep_key)
  tibble(
    n_rep_removed_vs_raw = nrow(removed),
    n_rep_added_vs_raw = nrow(added),
    n_rep_swaps_vs_raw = nrow(removed) + nrow(added)
  )
}

rep_identity_drift <- purrr::imap_dfr(
  dedup_results,
  function(scope_list, dataset_name) {
    purrr::imap_dfr(
      scope_list,
      function(res, scope_name) {
        raw_res <- dedup_results$sir_wide_raw[[scope_name]]
        drift <- compute_rep_identity_drift(raw_res, res, rep_key = rep_key)
        tibble(dataset = dataset_name, scope = scope_name) %>% bind_cols(drift)
      }
    )
  }
)

dedup_rep_drift_summary <- rep_identity_drift %>%
  left_join(
    dedup_summary %>% select(dataset, scope, n_retained_rows, n_within_class_discord_pairs),
    by = c("dataset", "scope")
  ) %>%
  mutate(
    pct_rep_swaps_vs_raw = ifelse(
      n_retained_rows > 0L,
      100 * n_rep_swaps_vs_raw / n_retained_rows,
      NA_real_
    )
  )

raw_dedup_baseline <- dedup_summary %>%
  filter(dataset == "sir_wide_raw") %>%
  select(
    scope,
    raw_n_retained_rows = n_retained_rows,
    raw_pct_episodes_multi_class = pct_episodes_multi_class,
    raw_order_sensitive_rate_pct = order_sensitive_rate_pct
  )

dedup_impact_summary <- dedup_summary %>%
  left_join(raw_dedup_baseline, by = "scope") %>%
  transmute(
    dataset,
    scope,
    delta_retained_rows = n_retained_rows - raw_n_retained_rows,
    delta_pct_episodes_multi_class = pct_episodes_multi_class - raw_pct_episodes_multi_class,
    delta_order_sensitive_rate_pct = order_sensitive_rate_pct - raw_order_sensitive_rate_pct
  ) %>%
  left_join(
    dedup_rep_drift_summary %>%
      select(
        dataset, scope,
        n_within_class_discord_pairs,
        n_rep_removed_vs_raw,
        n_rep_added_vs_raw,
        n_rep_swaps_vs_raw,
        pct_rep_swaps_vs_raw
      ),
    by = c("dataset", "scope")
  )

stopifnot(all(dedup_impact_summary$n_rep_swaps_vs_raw == dedup_impact_summary$n_rep_removed_vs_raw + dedup_impact_summary$n_rep_added_vs_raw))
if (any(dedup_impact_summary$delta_retained_rows == 0L)) {
  stopifnot(all(
    dedup_impact_summary$n_rep_removed_vs_raw[dedup_impact_summary$delta_retained_rows == 0L] ==
      dedup_impact_summary$n_rep_added_vs_raw[dedup_impact_summary$delta_retained_rows == 0L]
  ))
}

show_audit_table(
  data = dedup_summary,
  file_stem = "dedup_summary",
  caption = "Dedup summary",
  page_length = 20,
  scroll_y = "360px",
  format_round_cols = c("pct_episodes_multi_class", "order_sensitive_rate_pct", "pct_rep_swaps_vs_raw"),
  digits = 3
)

show_audit_table(
  data = dedup_impact_summary,
  file_stem = "dedup_impact_summary",
  caption = "Dedup impact summary vs raw",
  page_length = 20,
  scroll_y = "360px",
  format_round_cols = c(
    "delta_pct_episodes_multi_class",
    "delta_order_sensitive_rate_pct",
    "pct_rep_swaps_vs_raw"
  ),
  digits = 3
)

invisible(list(
  dedup_summary = dedup_summary,
  dedup_impact_summary = dedup_impact_summary
))
```

Decision: dedup strategy sensitivity is material when retained-row deltas or representative identity drift in `dedup_impact_summary` are non-trivial versus `sir_wide_raw`.

### Class partition drift

Question: beyond retained-row deltas, do completion strategies change class partitions inside dedup episodes?

Evidence: `partition_drift_summary` quantifies episode-level class partition drift vs raw for each scope.

```{r}
#| label: dedup_partition_drift_audit
episode_partition_signature <- function(class_map, scope_keys) {
  stopifnot(all(c(scope_keys, "phenotype_class") %in% names(class_map)))

  row_key_pool <- c(
    "PATID", "EVTID", "ELTID", "souche_id",
    "DATEPRELEV", "HEUREPRELEV", "naturepvt_norm", "bact_norm"
  )
  row_key_cols <- intersect(row_key_pool, names(class_map))
  if (length(row_key_cols) == 0L) {
    stop("No usable row-key columns available to build partition signature.", call. = FALSE)
  }

  dup_group_cols <- unique(c(scope_keys, row_key_cols))

  class_map %>%
    mutate(across(all_of(row_key_cols), as.character)) %>%
    group_by(across(all_of(dup_group_cols))) %>%
    mutate(.dup_idx = row_number()) %>%
    ungroup() %>%
    tidyr::unite("row_key", all_of(c(row_key_cols, ".dup_idx")), sep = "|", remove = FALSE, na.rm = FALSE) %>%
    group_by(across(all_of(c(scope_keys, "phenotype_class")))) %>%
    summarise(class_sig = paste(sort(row_key), collapse = "||"), .groups = "drop") %>%
    group_by(across(all_of(scope_keys))) %>%
    summarise(
      n_classes = n(),
      partition_sig = paste(sort(class_sig), collapse = "###"),
      .groups = "drop"
    )
}

compare_scope <- function(scope_name) {
  scope_keys <- dedup_scope_defs[[scope_name]]
  raw_sig <- episode_partition_signature(
    dedup_results$sir_wide_raw[[scope_name]]$class_map,
    scope_keys = scope_keys
  )

  purrr::imap_dfr(
    dedup_results,
    function(scope_list, dataset_name) {
      cur_sig <- episode_partition_signature(
        scope_list[[scope_name]]$class_map,
        scope_keys = scope_keys
      )

      raw_sig %>%
        full_join(cur_sig, by = scope_keys, suffix = c("_raw", "_cur")) %>%
        mutate(
          changed_n_classes = dplyr::coalesce(n_classes_raw, -1L) != dplyr::coalesce(n_classes_cur, -1L),
          changed_partition = dplyr::coalesce(partition_sig_raw, "__missing__") != dplyr::coalesce(partition_sig_cur, "__missing__")
        ) %>%
        summarise(
          dataset = dataset_name,
          scope = scope_name,
          n_episodes = n(),
          n_nclass_changed = sum(changed_n_classes),
          n_partition_changed = sum(changed_partition)
        )
    }
  )
}

partition_drift_summary <- bind_rows(
  compare_scope("by_type"),
  compare_scope("global")
) %>%
  arrange(scope, dataset)

stopifnot(all(
  partition_drift_summary %>%
    filter(dataset == "sir_wide_raw") %>%
    pull(n_partition_changed) == 0L
))
stopifnot(all(
  partition_drift_summary %>%
    filter(dataset == "sir_wide_raw") %>%
    pull(n_nclass_changed) == 0L
))

show_audit_table(
  data = partition_drift_summary,
  file_stem = "partition_drift_summary",
  caption = "Partition drift summary",
  page_length = 20,
  scroll_y = "340px"
)
```

```{r}
#| label: dedup_partition_drift_episodes
build_partition_drift_episodes <- function(scope_name, dataset_name) {
  scope_keys <- dedup_scope_defs[[scope_name]]
  raw_sig <- episode_partition_signature(
    dedup_results$sir_wide_raw[[scope_name]]$class_map,
    scope_keys = scope_keys
  )
  cur_sig <- episode_partition_signature(
    dedup_results[[dataset_name]][[scope_name]]$class_map,
    scope_keys = scope_keys
  )
 
  raw_sig %>%
    full_join(cur_sig, by = scope_keys, suffix = c("_raw", "_cur")) %>%
    mutate(
      changed_n_classes = dplyr::coalesce(n_classes_raw, -1L) != dplyr::coalesce(n_classes_cur, -1L),
      changed_partition = dplyr::coalesce(partition_sig_raw, "__missing__") != dplyr::coalesce(partition_sig_cur, "__missing__")
    ) %>%
    filter(changed_n_classes | changed_partition) %>%
    transmute(
      dataset = dataset_name,
      scope = scope_name,
      across(all_of(scope_keys)),
      n_classes_raw,
      n_classes_cur,
      changed_n_classes,
      changed_partition
    )
}

partition_drift_episodes <- purrr::map_dfr(
  setdiff(names(dedup_results), "sir_wide_raw"),
  function(dataset_name) {
    purrr::map_dfr(
      names(dedup_scope_defs),
      ~ build_partition_drift_episodes(scope_name = .x, dataset_name = dataset_name)
    )
  }
) %>%
  arrange(desc(changed_n_classes), dataset, scope)

show_audit_table(
  data = partition_drift_episodes,
  file_stem = "partition_drift_episodes",
  caption = "Partition drift episodes (changed partitions/classes)",
  page_length = 25,
  scroll_y = "500px"
)
```

### Swap decomposition

Question: are representative swaps mostly driven by partition changes or by representative re-ranking inside unchanged partitions?

Evidence: `swap_decomposition_summary` decomposes representative swaps (`removed + added`) by partition status.

```{r}
#| label: dedup_swap_decomposition_audit

swap_rep_key <- c(
  "PATID", "EVTID", "ELTID", "souche_id",
  "naturepvt_norm", "bact_norm", "DATEPRELEV", "HEUREPRELEV"
)

episode_partition_status <- function(scope_name, dataset_name) {
  scope_keys <- dedup_scope_defs[[scope_name]]
  raw_sig <- episode_partition_signature(
    dedup_results$sir_wide_raw[[scope_name]]$class_map,
    scope_keys = scope_keys
  )
  cur_sig <- episode_partition_signature(
    dedup_results[[dataset_name]][[scope_name]]$class_map,
    scope_keys = scope_keys
  )

  raw_sig %>%
    full_join(cur_sig, by = scope_keys, suffix = c("_raw", "_cur")) %>%
    mutate(
      changed_n_classes = dplyr::coalesce(n_classes_raw, -1L) != dplyr::coalesce(n_classes_cur, -1L),
      changed_partition = dplyr::coalesce(partition_sig_raw, "__missing__") != dplyr::coalesce(partition_sig_cur, "__missing__")
    ) %>%
    select(any_of(scope_keys), changed_n_classes, changed_partition)
}

rep_swap_events <- function(scope_name, dataset_name) {
  scope_keys <- dedup_scope_defs[[scope_name]]
  raw_res <- dedup_results$sir_wide_raw[[scope_name]]
  cur_res <- dedup_results[[dataset_name]][[scope_name]]

  key_cols <- intersect(
    swap_rep_key,
    intersect(names(raw_res$dedup), names(cur_res$dedup))
  )
  stopifnot(length(key_cols) > 0L)

  a <- raw_res$dedup %>%
    select(any_of(c(scope_keys, key_cols))) %>%
    distinct()
  b <- cur_res$dedup %>%
    select(any_of(c(scope_keys, key_cols))) %>%
    distinct()

  removed <- anti_join(a, b, by = key_cols) %>%
    mutate(swap_direction = "removed")
  added <- anti_join(b, a, by = key_cols) %>%
    mutate(swap_direction = "added")

  bind_rows(removed, added) %>%
    left_join(
      episode_partition_status(scope_name, dataset_name),
      by = scope_keys
    ) %>%
    mutate(
      dataset = dataset_name,
      scope = scope_name
    )
}

swap_decomposition_events <- purrr::map_dfr(
  setdiff(names(dedup_results), "sir_wide_raw"),
  function(dataset_name) {
    purrr::map_dfr(
      names(dedup_scope_defs),
      ~ rep_swap_events(scope_name = .x, dataset_name = dataset_name)
    )
  }
)

swap_decomposition_summary <- tidyr::expand_grid(
  dataset = names(dedup_results),
  scope = names(dedup_scope_defs)
) %>%
  left_join(
    swap_decomposition_events %>%
      group_by(dataset, scope) %>%
      summarise(
        n_rep_removed_vs_raw = sum(swap_direction == "removed"),
        n_rep_added_vs_raw = sum(swap_direction == "added"),
        n_rep_swaps_vs_raw = n(),
        n_rep_swaps_changed_partition = sum(changed_partition %in% TRUE, na.rm = TRUE),
        n_rep_swaps_same_partition = sum(changed_partition %in% FALSE, na.rm = TRUE),
        n_rep_swaps_nclass_changed = sum(changed_n_classes %in% TRUE, na.rm = TRUE),
        .groups = "drop"
      ),
    by = c("dataset", "scope")
  ) %>%
  mutate(
    across(
      c(
        n_rep_removed_vs_raw,
        n_rep_added_vs_raw,
        n_rep_swaps_vs_raw,
        n_rep_swaps_changed_partition,
        n_rep_swaps_same_partition,
        n_rep_swaps_nclass_changed
      ),
      ~ tidyr::replace_na(.x, 0L)
    ),
    pct_rep_swaps_changed_partition = ifelse(
      n_rep_swaps_vs_raw > 0L,
      100 * n_rep_swaps_changed_partition / n_rep_swaps_vs_raw,
      NA_real_
    ),
    pct_rep_swaps_same_partition = ifelse(
      n_rep_swaps_vs_raw > 0L,
      100 * n_rep_swaps_same_partition / n_rep_swaps_vs_raw,
      NA_real_
    )
  ) %>%
  arrange(scope, dataset)

swap_decomposition_episodes <- swap_decomposition_events %>%
  group_by(
    dataset, scope,
    PATID, dedup_year, naturepvt_norm, bact_norm,
    changed_n_classes, changed_partition
  ) %>%
  summarise(
    n_swap_rows = n(),
    n_removed = sum(swap_direction == "removed"),
    n_added = sum(swap_direction == "added"),
    .groups = "drop"
  ) %>%
  arrange(desc(n_swap_rows), desc(changed_partition), desc(changed_n_classes), dataset, scope)

stopifnot(all(swap_decomposition_summary$n_rep_swaps_vs_raw == swap_decomposition_summary$n_rep_removed_vs_raw + swap_decomposition_summary$n_rep_added_vs_raw))
stopifnot(all(swap_decomposition_summary$n_rep_swaps_vs_raw == swap_decomposition_summary$n_rep_swaps_changed_partition + swap_decomposition_summary$n_rep_swaps_same_partition))

swap_vs_dedup_summary_check <- dedup_impact_summary %>%
  select(dataset, scope, n_rep_swaps_vs_raw) %>%
  rename(n_rep_swaps_vs_raw_dedup = n_rep_swaps_vs_raw) %>%
  left_join(
    swap_decomposition_summary %>%
      select(dataset, scope, n_rep_swaps_vs_raw) %>%
      rename(n_rep_swaps_vs_raw_decomp = n_rep_swaps_vs_raw),
    by = c("dataset", "scope")
  )

stopifnot(all(swap_vs_dedup_summary_check$n_rep_swaps_vs_raw_dedup == swap_vs_dedup_summary_check$n_rep_swaps_vs_raw_decomp))

show_audit_table(
  data = swap_decomposition_summary,
  file_stem = "swap_decomposition_summary",
  caption = "Swap decomposition summary",
  page_length = 20,
  scroll_y = "360px",
  format_round_cols = c("pct_rep_swaps_changed_partition", "pct_rep_swaps_same_partition"),
  digits = 3
)
```

```{r}
#| label: dedup_swap_decomposition_episodes
show_audit_table(
  data = swap_decomposition_episodes,
  file_stem = "swap_decomposition_episodes",
  caption = "Swap decomposition episodes",
  page_length = 25,
  scroll_y = "500px"
)
```

Decision: completion strategy materially changes class structure when `partition_drift_summary$n_partition_changed > 0`, even if retained-row deltas are small.

Interpretation guide:

- if `n_rep_swaps_changed_partition` is dominant, completion is changing class boundaries in affected episodes.
- if `n_rep_swaps_same_partition` is dominant, completion mostly changes representative ranking inside stable partitions. **This seems to be the case.**
- if `delta_retained_rows` is near zero but `n_rep_swaps_vs_raw > 0`, dedup count stability hides representative identity drift.

### C3G indicator drift

Question: does completion strategy materially change C3G resistance indicators by `(year, naturepvt_norm, bact_norm)` after dedup?

Evidence: `c3g_indicator_impact_summary` compares dedup-level C3G counts/rates against raw baseline for each scope.

Column dictionary (source of truth):

- `c3g_indicator_panel` grain: one row per `(dataset, scope, dedup_year, naturepvt_norm, bact_norm)`.
- `dataset`: completion strategy (`sir_wide_raw`, `completion_date`, `completion_evtid`, `completion_patid_30d`, `completion_patid_year`).
- `scope`: dedup scope (`global` or `by_type`).
- `dedup_year`, `naturepvt_norm`, `bact_norm`: analysis keys.
- `n_isolates`: number of deduplicated isolates in the cell.
- `n_c3g_tested`: isolates with at least one informative C3G result (`S` or `R`) among `c3g_cols`.
- `n_c3g_resistant`: isolates with at least one resistant C3G result (`R`) among `c3g_cols`.
- `pct_c3g_resistant_all`: `100 * n_c3g_resistant / n_isolates`.
- `pct_c3g_resistant_tested`: `100 * n_c3g_resistant / n_c3g_tested`.

`c3g_indicator_impact` (cell-level drift vs raw baseline, same grain keys):

- Join logic: full join between raw and strategy panels by `(dedup_year, naturepvt_norm, bact_norm)` within each scope.
- Missing count columns after join are coerced to `0` (`*_raw`, `*_new` counts only).
- `dataset`, `scope`: strategy and scope of comparison.
- Raw/New count columns:
  - `n_isolates_raw`, `n_isolates_new`
  - `n_c3g_tested_raw`, `n_c3g_tested_new`
  - `n_c3g_resistant_raw`, `n_c3g_resistant_new`
- Raw/New rate columns (from panel):
  - `pct_c3g_resistant_all_raw`, `pct_c3g_resistant_all_new`
  - `pct_c3g_resistant_tested_raw`, `pct_c3g_resistant_tested_new`
- Delta count columns:
  - `delta_n_isolates = n_isolates_new - n_isolates_raw`
  - `delta_n_c3g_tested = n_c3g_tested_new - n_c3g_tested_raw`
  - `delta_n_c3g_resistant = n_c3g_resistant_new - n_c3g_resistant_raw`
- Delta rate columns:
  - `delta_pct_c3g_resistant_all = pct_c3g_resistant_all_new - pct_c3g_resistant_all_raw` only when both `n_isolates_raw > 0` and `n_isolates_new > 0`, else `NA`.
  - `delta_pct_c3g_resistant_tested = pct_c3g_resistant_tested_new - pct_c3g_resistant_tested_raw` only when both `n_c3g_tested_raw > 0` and `n_c3g_tested_new > 0`, else `NA`.

`c3g_indicator_impact_summary` (strategy/scope aggregate):

- `n_cells`: number of compared cells.
- `n_cells_pct_tested_comparable`, `n_cells_pct_tested_non_comparable`: count of cells with/without defined `delta_pct_c3g_resistant_tested`.
- `n_cells_pct_all_comparable`, `n_cells_pct_all_non_comparable`: count of cells with/without defined `delta_pct_c3g_resistant_all`.
- `n_cells_delta_n_resistant`: cells where `delta_n_c3g_resistant != 0`.
- `n_cells_delta_pct_tested`: cells where `abs(delta_pct_c3g_resistant_tested) > 0`.
- `max_abs_delta_n_resistant`: maximum absolute resistant-count drift across cells.
- `p95_abs_delta_pct_tested`: 95th percentile of `abs(delta_pct_c3g_resistant_tested)` on comparable cells.
- `total_delta_n_isolates`: net isolate-count drift across all cells.
- `total_delta_n_resistant`: net resistant-count drift across all cells.

Interpretation notes:

- Positive `delta_*` means strategy > raw baseline; negative means strategy < raw.
- `delta_pct_*` can be `NA` because denominator comparability is intentionally enforced (not an error).
- Invariant checks below enforce `total_delta_n_isolates == 0` and `total_delta_n_resistant == 0` for `dataset == \"sir_wide_raw\"`.

```{r}
#| label: dedup_c3g_indicator_impact

c3g_candidates <- c("ceftriaxone", "cefotaxime", "ceftazidime")
c3g_cols <- intersect(c3g_candidates, atb_cols)
if (length(c3g_cols) == 0L) {
  stop(
    "No C3G columns found in atb_cols. Expected at least one of: ceftriaxone, cefotaxime, ceftazidime.",
    call. = FALSE
  )
}

compute_c3g_panel <- function(ds, dataset_name, scope_name, c3g_cols) {
  stopifnot(is.data.frame(ds), "dedup_year" %in% names(ds))
  if (!"naturepvt_norm" %in% names(ds)) ds$naturepvt_norm <- NA_character_
  if (!"bact_norm" %in% names(ds)) ds$bact_norm <- NA_character_

  ds %>%
    mutate(
      n_c3g_tested_drugs = rowSums(across(all_of(c3g_cols), ~ .x %in% c("S", "R")), na.rm = TRUE),
      n_c3g_r_drugs = rowSums(across(all_of(c3g_cols), ~ .x == "R"), na.rm = TRUE),
      is_c3g_tested = n_c3g_tested_drugs > 0L,
      is_c3g_resistant = n_c3g_r_drugs > 0L
    ) %>%
    group_by(dedup_year, naturepvt_norm, bact_norm) %>%
    summarise(
      n_isolates = n(),
      n_c3g_tested = sum(is_c3g_tested, na.rm = TRUE),
      n_c3g_resistant = sum(is_c3g_resistant, na.rm = TRUE),
      pct_c3g_resistant_all = ifelse(n_isolates > 0L, 100 * n_c3g_resistant / n_isolates, NA_real_),
      pct_c3g_resistant_tested = ifelse(n_c3g_tested > 0L, 100 * n_c3g_resistant / n_c3g_tested, NA_real_),
      .groups = "drop"
    ) %>%
    mutate(dataset = dataset_name, scope = scope_name, .before = 1)
}

c3g_indicator_panel <- purrr::imap_dfr(
  dedup_results,
  function(scope_list, dataset_name) {
    purrr::imap_dfr(
      scope_list,
      function(res, scope_name) {
        compute_c3g_panel(
          ds = res$dedup,
          dataset_name = dataset_name,
          scope_name = scope_name,
          c3g_cols = c3g_cols
        )
      }
    )
  }
)

impact_keys <- c("dedup_year", "naturepvt_norm", "bact_norm")

compute_c3g_impact_vs_raw <- function(dataset_name, scope_name) {
  raw_tbl <- c3g_indicator_panel %>%
    filter(dataset == "sir_wide_raw", scope == scope_name) %>%
    select(-dataset, -scope)
  new_tbl <- c3g_indicator_panel %>%
    filter(dataset == dataset_name, scope == scope_name) %>%
    select(-dataset, -scope)

  full_join(raw_tbl, new_tbl, by = impact_keys, suffix = c("_raw", "_new")) %>%
    mutate(
      dataset = dataset_name,
      scope = scope_name,
      n_isolates_raw = tidyr::replace_na(n_isolates_raw, 0L),
      n_c3g_tested_raw = tidyr::replace_na(n_c3g_tested_raw, 0L),
      n_c3g_resistant_raw = tidyr::replace_na(n_c3g_resistant_raw, 0L),
      n_isolates_new = tidyr::replace_na(n_isolates_new, 0L),
      n_c3g_tested_new = tidyr::replace_na(n_c3g_tested_new, 0L),
      n_c3g_resistant_new = tidyr::replace_na(n_c3g_resistant_new, 0L),
      delta_n_isolates = n_isolates_new - n_isolates_raw,
      delta_n_c3g_tested = n_c3g_tested_new - n_c3g_tested_raw,
      delta_n_c3g_resistant = n_c3g_resistant_new - n_c3g_resistant_raw,
      delta_pct_c3g_resistant_all = dplyr::if_else(
        n_isolates_raw > 0L & n_isolates_new > 0L,
        pct_c3g_resistant_all_new - pct_c3g_resistant_all_raw,
        NA_real_
      ),
      delta_pct_c3g_resistant_tested = dplyr::if_else(
        n_c3g_tested_raw > 0L & n_c3g_tested_new > 0L,
        pct_c3g_resistant_tested_new - pct_c3g_resistant_tested_raw,
        NA_real_
      )
    )
}

c3g_indicator_impact <- purrr::map_dfr(
  names(dedup_results),
  function(dataset_name) {
    purrr::map_dfr(
      names(dedup_scope_defs),
      ~ compute_c3g_impact_vs_raw(dataset_name = dataset_name, scope_name = .x)
    )
  }
) %>%
  relocate(dataset, scope, all_of(impact_keys))

c3g_indicator_impact_summary <- c3g_indicator_impact %>%
  group_by(dataset, scope) %>%
  summarise(
    n_cells = n(),
    n_cells_pct_tested_comparable = sum(!is.na(delta_pct_c3g_resistant_tested)),
    n_cells_pct_tested_non_comparable = sum(is.na(delta_pct_c3g_resistant_tested)),
    n_cells_pct_all_comparable = sum(!is.na(delta_pct_c3g_resistant_all)),
    n_cells_pct_all_non_comparable = sum(is.na(delta_pct_c3g_resistant_all)),
    n_cells_delta_n_resistant = sum(delta_n_c3g_resistant != 0L, na.rm = TRUE),
    n_cells_delta_pct_tested = sum(abs(delta_pct_c3g_resistant_tested) > 0, na.rm = TRUE),
    max_abs_delta_n_resistant = max(abs(delta_n_c3g_resistant), na.rm = TRUE),
    p95_abs_delta_pct_tested = if (all(is.na(delta_pct_c3g_resistant_tested))) {
      NA_real_
    } else {
      suppressWarnings(as.numeric(quantile(abs(delta_pct_c3g_resistant_tested), 0.95, na.rm = TRUE)))
    },
    total_delta_n_isolates = sum(delta_n_isolates, na.rm = TRUE),
    total_delta_n_resistant = sum(delta_n_c3g_resistant, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(scope, dataset)

stopifnot(all(
  c3g_indicator_impact_summary %>%
    filter(dataset == "sir_wide_raw") %>%
    pull(total_delta_n_isolates) == 0L
))
stopifnot(all(
  c3g_indicator_impact_summary %>%
    filter(dataset == "sir_wide_raw") %>%
    pull(total_delta_n_resistant) == 0L
))

show_audit_table(
  data = c3g_indicator_impact_summary,
  file_stem = "c3g_indicator_impact_summary",
  caption = "C3G indicator impact summary",
  page_length = 20,
  scroll_y = "360px",
  digits = datatable_digits
)
```

```{r}
#| label: dedup_c3g_indicator_top_changes
c3g_indicator_top_changes <- c3g_indicator_impact %>%
  filter(dataset != "sir_wide_raw") %>%
  mutate(
    abs_delta_n_c3g_resistant = abs(delta_n_c3g_resistant),
    abs_delta_pct_c3g_resistant_tested = abs(delta_pct_c3g_resistant_tested)
  ) %>%
  group_by(dataset, scope) %>%
  slice_max(order_by = abs_delta_n_c3g_resistant, n = 20, with_ties = FALSE) %>%
  ungroup() %>%
  arrange(desc(abs_delta_n_c3g_resistant), desc(abs_delta_pct_c3g_resistant_tested), dataset, scope)

show_audit_table(
  data = c3g_indicator_top_changes,
  file_stem = "c3g_indicator_top_changes",
  caption = "C3G top changes (top 20 by dataset/scope)",
  page_length = 25,
  scroll_y = "500px",
  digits = datatable_digits
)
```

```{r}
#| label: dedup_c3g_indicator_decision_metrics

c3g_indicator_decision_metrics <- c3g_indicator_impact %>%
  filter(dataset != "sir_wide_raw") %>%
  summarise(
    max_abs_delta_n_resistant = max(abs(delta_n_c3g_resistant), na.rm = TRUE),
    n_cells_pct_tested_comparable = sum(!is.na(delta_pct_c3g_resistant_tested)),
    p95_abs_delta_pct_tested = if (all(is.na(delta_pct_c3g_resistant_tested))) {
      NA_real_
    } else {
      suppressWarnings(as.numeric(quantile(abs(delta_pct_c3g_resistant_tested), 0.95, na.rm = TRUE)))
    },
    n_positive_delta = sum(delta_n_c3g_resistant > 0, na.rm = TRUE),
    n_negative_delta = sum(delta_n_c3g_resistant < 0, na.rm = TRUE),
    n_zero_delta = sum(delta_n_c3g_resistant == 0, na.rm = TRUE)
  )

show_audit_table(
  data = c3g_indicator_decision_metrics,
  file_stem = "c3g_indicator_decision_metrics",
  caption = "C3G decision metrics",
  page_length = 10,
  scroll_y = "260px",
  digits = datatable_digits
)
```

Decision: C3G drift is interpreted using magnitude (`max_abs_delta_n_resistant`) and spread (`p95_abs_delta_pct_tested`) on comparable denominator cells only, with directionality checked through positive/negative delta counts. In this run, `max_abs_delta_n_resistant = `r c3g_indicator_decision_metrics$max_abs_delta_n_resistant`, `p95_abs_delta_pct_tested = `r round(c3g_indicator_decision_metrics$p95_abs_delta_pct_tested, 3)`, `n_cells_pct_tested_comparable = `r c3g_indicator_decision_metrics$n_cells_pct_tested_comparable`, and directionality is mixed (`+`: `r c3g_indicator_decision_metrics$n_positive_delta`, `-`: `r c3g_indicator_decision_metrics$n_negative_delta`).

### C3G operational rates

```{r}
#| label: dedup_c3g_rates_all_datasets

compute_c3g_rate <- function(ds, group_vars, c3g_cols) {
  stopifnot(all(group_vars %in% names(ds)))
  stopifnot(length(c3g_cols) > 0L, all(c3g_cols %in% names(ds)))

  ds %>%
    mutate(
      n_c3g_tested_drugs = rowSums(across(all_of(c3g_cols), ~ .x %in% c("S", "R")), na.rm = TRUE),
      n_c3g_r_drugs = rowSums(across(all_of(c3g_cols), ~ .x == "R"), na.rm = TRUE),
      is_c3g_tested = n_c3g_tested_drugs > 0L,
      is_c3g_resistant = n_c3g_r_drugs > 0L
    ) %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(
      n_tested = sum(is_c3g_tested, na.rm = TRUE),
      n_resistant = sum(is_c3g_resistant, na.rm = TRUE),
      taux_resistance_c3g = ifelse(n_tested > 0L, 100 * n_resistant / n_tested, NA_real_),
      .groups = "drop"
    ) %>%
    arrange(across(all_of(group_vars)))
}

compute_c3g_rate_by_scope <- function(scope_name, group_vars, c3g_cols) {
  purrr::imap_dfr(
    dedup_results,
    function(scope_list, dataset_name) {
      compute_c3g_rate(
        ds = scope_list[[scope_name]]$dedup,
        group_vars = group_vars,
        c3g_cols = c3g_cols
      ) %>%
        mutate(dataset = dataset_name, scope = scope_name, .before = 1)
    }
  )
}

c3g_rate_global_dedup <- compute_c3g_rate_by_scope(
  scope_name = "global",
  group_vars = c("dedup_year", "bact_norm"),
  c3g_cols = c3g_cols
)

c3g_rate_by_type_dedup <- compute_c3g_rate_by_scope(
  scope_name = "by_type",
  group_vars = c("dedup_year", "naturepvt_norm", "bact_norm"),
  c3g_cols = c3g_cols
)

compute_c3g_rate_drift_vs_raw <- function(rate_tbl, join_keys) {
  raw_tbl <- rate_tbl %>%
    filter(dataset == "sir_wide_raw") %>%
    select(any_of(c("scope", join_keys, "n_tested", "n_resistant", "taux_resistance_c3g"))) %>%
    rename(
      n_tested_raw = n_tested,
      n_resistant_raw = n_resistant,
      taux_resistance_c3g_raw = taux_resistance_c3g
    )

  datasets <- setdiff(unique(rate_tbl$dataset), "sir_wide_raw")

  purrr::map_dfr(
    datasets,
    function(dataset_name) {
      new_tbl <- rate_tbl %>%
        filter(dataset == dataset_name) %>%
        select(any_of(c("scope", join_keys, "n_tested", "n_resistant", "taux_resistance_c3g"))) %>%
        rename(
          n_tested_new = n_tested,
          n_resistant_new = n_resistant,
          taux_resistance_c3g_new = taux_resistance_c3g
        )

      full_join(new_tbl, raw_tbl, by = c("scope", join_keys)) %>%
        mutate(
          dataset = dataset_name,
          n_tested_raw = tidyr::replace_na(n_tested_raw, 0L),
          n_resistant_raw = tidyr::replace_na(n_resistant_raw, 0L),
          n_tested_new = tidyr::replace_na(n_tested_new, 0L),
          n_resistant_new = tidyr::replace_na(n_resistant_new, 0L),
          delta_n_tested = n_tested_new - n_tested_raw,
          delta_n_resistant = n_resistant_new - n_resistant_raw,
          delta_taux_resistance_c3g = dplyr::if_else(
            n_tested_raw > 0L & n_tested_new > 0L,
            taux_resistance_c3g_new - taux_resistance_c3g_raw,
            NA_real_
          ),
          abs_delta_n_resistant = abs(delta_n_resistant),
          abs_delta_taux_resistance_c3g = abs(delta_taux_resistance_c3g)
        )
    }
  ) %>%
    arrange(desc(abs_delta_n_resistant), desc(abs_delta_taux_resistance_c3g), dataset)
}

c3g_rate_global_drift_vs_raw <- compute_c3g_rate_drift_vs_raw(
  rate_tbl = c3g_rate_global_dedup,
  join_keys = c("dedup_year", "bact_norm")
)

c3g_rate_by_type_drift_vs_raw <- compute_c3g_rate_drift_vs_raw(
  rate_tbl = c3g_rate_by_type_dedup,
  join_keys = c("dedup_year", "naturepvt_norm", "bact_norm")
)

# Compatibility shims: keep raw-only aliases for older downstream snippets.
c3g_rate_global_raw_dedup <- c3g_rate_global_dedup %>%
  filter(dataset == "sir_wide_raw") %>%
  select(-dataset, -scope)

c3g_rate_by_type_raw_dedup <- c3g_rate_by_type_dedup %>%
  filter(dataset == "sir_wide_raw") %>%
  select(-dataset, -scope)

c3g_rate_global_dedup <- c3g_rate_global_dedup %>% arrange(scope, dataset, desc(n_tested))
c3g_rate_by_type_dedup <- c3g_rate_by_type_dedup %>% arrange(scope, dataset, desc(n_tested))

show_audit_table(
  data = c3g_rate_global_dedup,
  file_stem = "c3g_rate_global_dedup",
  caption = "C3G rates global (all datasets)",
  page_length = 25,
  scroll_y = "500px",
  digits = datatable_digits
)

show_audit_table(
  data = c3g_rate_by_type_dedup,
  file_stem = "c3g_rate_by_type_dedup",
  caption = "C3G rates by type (all datasets)",
  page_length = 25,
  scroll_y = "500px",
  digits = datatable_digits
)

show_audit_table(
  data = c3g_rate_global_drift_vs_raw,
  file_stem = "c3g_rate_global_drift_vs_raw",
  caption = "C3G global drift vs raw",
  page_length = 25,
  scroll_y = "500px",
  digits = datatable_digits
)

show_audit_table(
  data = c3g_rate_by_type_drift_vs_raw,
  file_stem = "c3g_rate_by_type_drift_vs_raw",
  caption = "C3G by-type drift vs raw",
  page_length = 25,
  scroll_y = "500px",
  digits = datatable_digits
)

invisible(list(
  c3g_rate_global_dedup = c3g_rate_global_dedup,
  c3g_rate_by_type_dedup = c3g_rate_by_type_dedup,
  c3g_rate_global_drift_vs_raw = c3g_rate_global_drift_vs_raw,
  c3g_rate_by_type_drift_vs_raw = c3g_rate_by_type_drift_vs_raw
))
```

# Resistance indicators production

Product-facing RATB indicator production has moved to `orchidee_ratb_indicators.qmd`.

This split keeps the current notebook focused on:

- hospitalization scope construction
- completion strategies
- dedup behavior and diagnostics

and makes the rendered HTML materially lighter and faster to open.

Recommended render order:

1. `quarto render orchidee_dedup_workflow.qmd`
2. `quarto render orchidee_ratb_indicators.qmd`

## RATB QA appendix

This appendix centralizes the section-6 QA material that no longer belongs in
the product-facing indicators notebook.

### Indicator execution QA

```{r}
#| label: ratb_indicator_qa_build
#| echo: false
stopifnot(
  exists("load_ratb_indicator_spec", mode = "function"),
  exists("build_species_taxonomy_map", mode = "function"),
  exists("validate_ratb_indicator_spec", mode = "function"),
  exists("build_ratb_indicator_coverage_audit", mode = "function"),
  exists("build_ratb_indicator_panel_annual", mode = "function")
)

ratb_indicator_spec_path <- orchidee_config$paths$ratb_indicator_spec_path
ratb_indicator_datasets <- names(dedup_results)
species_taxonomy_path <- file.path(dictionaries_dir, "species_regex_map.csv")
bact_order_map <- build_species_taxonomy_map(species_taxonomy_path)

reference_dataset_name <- names(dedup_results)[[1]]
reference_global_dedup <- dedup_results[[reference_dataset_name]]$global$dedup
reference_by_type_dedup <- dedup_results[[reference_dataset_name]]$by_type$dedup

atb_cols <- intersect(sir_wide_meta$atb_cols, names(reference_global_dedup))
stopifnot(length(atb_cols) > 0L)

available_phenotype_cols <- intersect(
  c("blse_flag", "carbapenemase_flag"),
  names(reference_global_dedup)
)
available_sample_types <- sort(unique(na.omit(as.character(reference_by_type_dedup$naturepvt_norm))))

ratb_indicator_spec <- load_ratb_indicator_spec(ratb_indicator_spec_path)
ratb_indicator_spec_validation <- validate_ratb_indicator_spec(
  spec = ratb_indicator_spec,
  atb_cols = atb_cols,
  phenotype_cols = available_phenotype_cols,
  available_sample_types = available_sample_types
)

stopifnot(!any(ratb_indicator_spec_validation$duplicated_indicator_id))
stopifnot(all(ratb_indicator_spec_validation$supported_indicator_kind))
stopifnot(all(ratb_indicator_spec_validation$supported_sample_type_mode))
stopifnot(all(ratb_indicator_spec_validation$supported_numerator_kind))
stopifnot(all(ratb_indicator_spec_validation$supported_denominator_kind))
stopifnot(all(ratb_indicator_spec_validation$has_scope_mode))
stopifnot(all(ratb_indicator_spec_validation$has_filter_values))
stopifnot(all(ratb_indicator_spec_validation$has_valid_indicator_payload))

ratb_indicator_coverage_audit <- build_ratb_indicator_coverage_audit(
  spec = ratb_indicator_spec,
  validation = ratb_indicator_spec_validation
)

ratb_indicator_spec_wave1_exec <- ratb_indicator_spec %>%
  semi_join(
    ratb_indicator_coverage_audit %>%
      filter(executable_wave1) %>%
      select(indicator_id),
    by = "indicator_id"
  )

ratb_indicator_panel_annual <- build_ratb_indicator_panel_annual(
  dedup_results = dedup_results,
  spec = ratb_indicator_spec_wave1_exec,
  atb_cols = atb_cols,
  bact_order_map = bact_order_map
)

ratb_indicator_panel_global <- ratb_indicator_panel_annual %>%
  filter(scope == "global")
ratb_indicator_panel_by_type <- ratb_indicator_panel_annual %>%
  filter(scope == "by_type")

stopifnot(nrow(ratb_indicator_panel_annual) > 0L)
stopifnot(all(ratb_indicator_panel_annual$n_tested == ratb_indicator_panel_annual$n_r + ratb_indicator_panel_annual$n_s))
stopifnot(all(ratb_indicator_panel_annual$n_isolates == ratb_indicator_panel_annual$n_tested + ratb_indicator_panel_annual$n_o))
stopifnot(all(
  is.na(ratb_indicator_panel_annual$pct_resistant) |
    (ratb_indicator_panel_annual$pct_resistant >= 0 & ratb_indicator_panel_annual$pct_resistant <= 100)
))
```

```{r}
#| label: ratb_indicator_validation_table
show_audit_table(
  data = ratb_indicator_spec_validation %>%
    select(
      indicator_id,
      duplicated_indicator_id,
      supported_indicator_kind,
      supported_sample_type_mode,
      n_requested_molecules,
      n_available_molecules,
      all_required_molecules_available,
      phenotype_col_present,
      all_requested_sample_types_present,
      hard_validation_ok
    ),
  file_stem = "ratb_indicator_spec_validation",
  caption = "RATB indicator spec validation",
  page_length = 25,
  scroll_y = "420px"
)
```

```{r}
#| label: ratb_indicator_coverage_table
show_audit_table(
  data = ratb_indicator_coverage_audit %>%
    select(
      display_order, report_taxon_label, indicator_label,
      indicator_kind, publication_mode,
      proportion_execution_status, incidence_execution_status,
      coverage_note
    ),
  file_stem = "ratb_indicator_coverage_audit",
  caption = "RATB coverage audit",
  page_length = 25,
  scroll_y = "520px"
)
```

```{r}
#| label: ratb_indicator_overlap_regression
#| include: false
c3g_species_map <- tibble(
  report_taxon_label = c(
    "Escherichia coli",
    "Klebsiella pneumoniae",
    "Enterobacter cloacae complex"
  ),
  bact_norm = c(
    "escherichia_coli",
    "klebsiella_pneumoniae",
    "enterobacter_cloacae_complex"
  )
)

c3g_new_global <- ratb_indicator_panel_global %>%
  filter(indicator_label == "C3G") %>%
  inner_join(c3g_species_map, by = "report_taxon_label") %>%
  select(
    dataset, dedup_year, bact_norm,
    n_tested_new = n_tested,
    n_resistant_new = n_resistant,
    pct_resistant_new = pct_resistant
  )

c3g_old_global <- c3g_rate_global_dedup %>%
  filter(
    dataset %in% ratb_indicator_datasets,
    bact_norm %in% c3g_species_map$bact_norm
  ) %>%
  select(
    dataset, dedup_year, bact_norm,
    n_tested_old = n_tested,
    n_resistant_old = n_resistant,
    pct_resistant_old = taux_resistance_c3g
  )

c3g_global_consistency <- full_join(
  c3g_new_global,
  c3g_old_global,
  by = c("dataset", "dedup_year", "bact_norm")
) %>%
  mutate(
    delta_n_tested = n_tested_new - n_tested_old,
    delta_n_resistant = n_resistant_new - n_resistant_old,
    delta_pct_resistant = case_when(
      is.na(pct_resistant_new) & is.na(pct_resistant_old) ~ 0,
      TRUE ~ pct_resistant_new - pct_resistant_old
    )
  )

stopifnot(all(c3g_global_consistency$delta_n_tested == 0L))
stopifnot(all(c3g_global_consistency$delta_n_resistant == 0L))
stopifnot(all(abs(c3g_global_consistency$delta_pct_resistant) < 1e-8))

c3g_new_by_type <- ratb_indicator_panel_by_type %>%
  filter(indicator_label == "C3G") %>%
  inner_join(c3g_species_map, by = "report_taxon_label") %>%
  select(
    dataset, dedup_year, sample_type, bact_norm,
    n_tested_new = n_tested,
    n_resistant_new = n_resistant,
    pct_resistant_new = pct_resistant
  )

c3g_old_by_type <- c3g_rate_by_type_dedup %>%
  filter(
    dataset %in% ratb_indicator_datasets,
    bact_norm %in% c3g_species_map$bact_norm
  ) %>%
  select(
    dataset, dedup_year,
    sample_type = naturepvt_norm,
    bact_norm,
    n_tested_old = n_tested,
    n_resistant_old = n_resistant,
    pct_resistant_old = taux_resistance_c3g
  )

c3g_by_type_consistency <- full_join(
  c3g_new_by_type,
  c3g_old_by_type,
  by = c("dataset", "dedup_year", "sample_type", "bact_norm")
) %>%
  mutate(
    delta_n_tested = n_tested_new - n_tested_old,
    delta_n_resistant = n_resistant_new - n_resistant_old,
    delta_pct_resistant = case_when(
      is.na(pct_resistant_new) & is.na(pct_resistant_old) ~ 0,
      TRUE ~ pct_resistant_new - pct_resistant_old
    )
  )

stopifnot(all(c3g_by_type_consistency$delta_n_tested == 0L))
stopifnot(all(c3g_by_type_consistency$delta_n_resistant == 0L))
stopifnot(all(abs(c3g_by_type_consistency$delta_pct_resistant) < 1e-8))
```

### Denominator and perimeter QA

The denominator and perimeter work below now supports the published
incidence-density outputs. It remains in the backbone notebook because it is
still methodological QA and implementation documentation, not the primary
product view.

```{r}
#| label: ratb_perimeter_rules_table
show_audit_table(
  data = ratb_perimeter_rules,
  file_stem = "ratb_perimeter_rules",
  caption = "RATB TA/DE perimeter policy",
  page_length = 20,
  scroll_y = "280px"
)
```

```{r}
#| label: ratb_uf_ta_de_reference_table
show_audit_table(
  data = ratb_uf_ta_de_reference %>%
    select(
      SEJUF, consores_uf_label, CODE_TA, consores_ta_label,
      CODE_DE, de_domain_ref, consores_de_label, consores_care_type,
      uf_is_eligible_by_ta_de, uf_ta_de_status, uf_ta_de_reason
    ),
  file_stem = "ratb_uf_ta_de_reference",
  caption = "RATB UF TA/DE eligibility reference",
  page_length = 20,
  scroll_y = "360px"
)
```

```{r}
#| label: ratb_episode_exclusion_summary_table
show_audit_table(
  data = ratb_episode_exclusion_summary,
  file_stem = "ratb_episode_exclusion_summary",
  caption = "TA/DE incidence perimeter exclusion summary",
  page_length = 20,
  scroll_y = "320px",
  digits = datatable_digits
)
```

```{r}
#| label: hospital_days_year_summary_provisional_table
show_audit_table(
  data = hospital_days_year_summary_provisional,
  file_stem = "hospital_days_year_summary_provisional",
  caption = "Annual eligible hospital nights from PMSI TA/DE activity",
  page_length = 20,
  scroll_y = "260px",
  digits = datatable_digits
)
```

```{r}
#| label: ratb_numerator_scope_impact_audit_table
show_audit_table(
  data = ratb_numerator_scope_impact_audit,
  file_stem = "ratb_numerator_scope_impact_audit",
  caption = "Incidence numerator impact under the TA/DE perimeter",
  page_length = 20,
  scroll_y = "320px",
  digits = datatable_digits
)
```

```{r}
#| label: ratb_episode_scope_audit_table
show_audit_table(
  data = ratb_episode_scope_audit %>%
    filter(
      provisional_perimeter_status != "included" |
        zero_nights |
        cross_year
    ) %>%
    select(
      PATID, EVTID, provisional_perimeter_status, final_rule_level, final_reason,
      nights_provisional, zero_nights, cross_year, ratb_scope_status,
      eligible_uf_codes, noneligible_uf_codes, episode_ta_codes,
      episode_de_codes, episode_de_domains, episode_ta_de_reasons,
      episode_uf_codes, episode_uf_labels, episode_um_codes, episode_um_labels,
      episode_ghm_values
    ),
  file_stem = "ratb_episode_scope_audit",
  caption = "TA/DE incidence perimeter episode audit",
  page_length = 20,
  scroll_y = "360px",
  digits = datatable_digits
)
```

```{r}
#| label: hospital_stay_validation_summary_table
show_audit_table(
  data = hospital_stay_validation_summary,
  file_stem = "hospital_stay_validation_summary",
  caption = "Hospital stay validation summary",
  page_length = 20,
  scroll_y = "280px",
  digits = datatable_digits
)
```

```{r}
#| label: hospital_days_year_summary_table
show_audit_table(
  data = hospital_days_year_summary,
  file_stem = "hospital_days_year_summary",
  caption = "Hospital-days annual validation panel",
  page_length = 20,
  scroll_y = "320px",
  digits = datatable_digits
)
```

```{r}
#| label: hospital_stays_validation_audit_table
show_audit_table(
  data = hospital_stays_raw %>%
    filter(validation_status != "validated" | cross_year) %>%
    select(
      PATID, EVTID, validation_status, ratb_scope_status, datent_min, datsort_max,
      exact_elapsed_days, cross_year, evtid_multi_pat, pmsi_status_values,
      sejdur_values
    ),
  file_stem = "hospital_stays_validation_audit",
  caption = "Hospital stay audit (invalid or cross-year stays)",
  page_length = 20,
  scroll_y = "360px",
  digits = datatable_digits
)
```





















